---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# interflex

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-stable-green.svg)](https://www.tidyverse.org/lifecycle/#stablel)
[![License: MIT](https://img.shields.io/badge/License-MIT-yellow.svg)](https://opensource.org/licenses/MIT)
[![downloads: CRAN](https://cranlogs.r-pkg.org/badges/grand-total/interflex)](https://www.datasciencemeta.com/rpackages)
<!-- badges: end -->

**interflex** estimates, interprets, and visualizes marginal effects and performs diagnostics.

**Authors:** Jens Hainmueller, Jiehan Liu, Licheng Liu, Ziyi Liu, Jonathan Mummolo, Tianzhu Qin,  and [Yiqing Xu](https://yiqingxu.org/) (maintainer)

**Date:** June 18, 2024

**Repos:** [Github](https://github.com/xuyiqing/interflex) (1.3.0)
[CRAN](https://cran.r-project.org/web/packages/interflex/index.html) (1.2.6)

**Tutoralis:** See tutorials for cases with [continuous](https://yiqingxu.org/packages/interflex/articles/continuous.html) and [discrete](https://yiqingxu.org/packages/interflex/articles/discrete.html) outcomes, as well as a tutorial for double/debiased machine learning ([DML](https://yiqingxu.org/packages/interflex/articles/dml.html)) estimators. 

**Examples:** R code used in the tutorials can be downloaded from [here](examples.R).

**Reference:** [*How Much Should We Trust Estimates from Multiplicative Interaction Models? Simple Tools to Improve Empirical Practice*](http://bit.ly/HMX2019). Political Analysis, Vol. 27, Iss. 2, April 2019, pp. 163--192.

---

## Installation

You can install the **interflex** package from CRAN: 
```{r eval=FALSE}
install.packages('interflex', type = "source", repos = 'http://cran.us.r-project.org') 
```

Or you can install the up-to-date development version from Github:  
```{r eval=FALSE}
install.packages('devtools', repos = 'http://cran.us.r-project.org') # if not already installed
devtools::install_github('xuyiqing/interflex@dml')
```

**interflex** depends on the following packages, which will be installed automatically when **interflex** is being installed; if not, please install them manually:  
```{r echo=TRUE, message=FALSE, warning=FALSE}
# Function to install packages
install_all <- function(packages) {
  installed_pkgs <- installed.packages()[, "Package"]
  for (pkg in packages) {
    if (!pkg %in% installed_pkgs) {
      install.packages(pkg, repos = 'http://cran.us.r-project.org')
    }
  }
}

# Packages to be installed
packages <- c("Rcpp", "mgcv", "sandwich", "pcse", "foreach", "doParallel", 
              "lmtest", "lfe", "Lmoments", "ggplot2", "plotrix", "grid", 
              "gridExtra", "ggplotify", "ggpubr", "RColorBrewer", "grDevices", 
              "gtable", "MASS", "mvtnorm", "pROC", "ModelMetrics", "foreign",
              "patchwork", "rmarkdown")

# Install the packages
install_all(packages)
```
Mac users who encounter "-lgfortran" or "-lquadmath" error during installation, please check out the solution [here](http://thecoatlessprofessor.com/programming/rcpp-rcpparmadillo-and-os-x-mavericks-lgfortran-and-lquadmath-error/). Typing the following two lines of code in your **Terminal** should solve this problem.
```{r, eval = FALSE}
curl -OL http://r.research.att.com/libs/gfortran-4.8.2-darwin13.tar.bz2
sudo tar fvxz gfortran-4.8.2-darwin13.tar.bz2 -C /
```
Mac users who encounter **clang: error: unsupported option '-fopenmp'**, please consider (1) updating your R and/or (2) installing new R macro tools from [Github](https://github.com/rmacoslib/r-macos-rtools/releases/tag/v3.1.0).

## Python Environment

To use the double/debiased machine learning estimators in **interflex**, we rely on packages in Python. The integration between R and Python is facilitated by the **reticulate** package, which allows R to interface with Python seamlessly. The following steps set up a Python environment with all required dependencies.

First, install the **reticulate** package in R. This package enables R to interface with Python.

```{r eval=FALSE, include=TRUE}
install.packages("reticulate", repos = 'http://cran.us.r-project.org')
```

Then, we use [Miniconda](https://docs.anaconda.com/free/miniconda/), a minimal installer for Conda. Miniconda helps manage Python environments and packages. By default, the installation process will create a virtual environment named `r-reticulate`. 

```{r eval=FALSE, include=TRUE}
reticulate::install_miniconda(update = TRUE)
```

With Miniconda installed, we point the platform using `use_condaenv`. This step is incorporated in the **interflex** package.

```{r eval=FALSE, include=TRUE}
reticulate::use_condaenv(condaenv = "r-reticulate")
```

Finally, install the necessary Python libraries. These libraries include tools for statistical modeling and machine learning that are essential for using the interflex package.

```{r eval=FALSE, include=TRUE}
reticulate::py_install(packages = c("patsy", "numpy", "pandas", 
                    "scikit-learn==1.2.0", "doubleml", "econml"))
```


## Report bugs

Please report bugs to **yiqingxu [at] stanford.edu** with your sample code and data file. Much appreciated!


