# Get Started {#sec-introduction}

## Installation

To install **interflex** from CRAN, run the code chunk below:

```{r install-cran, eval=FALSE, message=FALSE, warning=FALSE}
install.packages("interflex")
```

We recommend users to install the most up-to-date, stable version of **interflex** from GitHub using:

```{r install-github-main, eval=FALSE, message=FALSE, warning=FALSE}
devtools::install_github("xuyiqing/interflex")
```

We fix bugs in the `dev` branch before merging into the main branch; therefore, it is often more up to date.

```{r install-github-dev, eval=FALSE, message=FALSE, warning=FALSE}
devtools::install_github("xuyiqing/interflex@dev")
```

After installation, check **interflex** version to make sure the package is up-to-date.

```{r ch1-check-version}
installed.packages()["interflex", "Version"]
```

**interflex** depends on the following packages, which should be installed automatically when **interflex** is being installed. You can also install them manually.

```{r install-dependencies, eval=FALSE}
install_all <- function(packages) {
  installed_pkgs <- installed.packages()[, "Package"]
  for (pkg in packages) {
    if (!pkg %in% installed_pkgs) {
      install.packages(pkg)
    }
  }
}
packages <- c("Rcpp", "RcppArmadillo", "ggplot2", "sandwich", "Lmoments",
              "doParallel", "foreach", "mgcv", "lfe", "gridExtra", "grid",
              "ggplotify", "fixest", "RColorBrewer", "pcse", "gtable",
              "MASS", "mvtnorm", "pROC", "ModelMetrics", "lmtest", "AER",
              "future", "glmnet", "grf", "DoubleML", "mlr3", "mlr3learners", 
              "data.table", "paradox", "ranger", "xgboost")
install_all(packages)
```

::: {.callout-note}
## xgboost and mlr3 compatibility

The DML estimator's histogram-based gradient boosting option (`model.y = "hgb"`) requires **xgboost**. Starting with **xgboost** ≥ 3.0, you need **mlr3** ≥ 1.6.0 and **mlr3learners** ≥ 0.14.0. If you encounter an error like `ALTLIST classes must provide a Set_elt method`, update both:

```r
install.packages(c("mlr3", "mlr3learners"))
```
:::


To use additional `dml` estimators, **interflex** suggests the additional **mlr3extralearners** package, which is not on CRAN but can be installed from GitHub:
```{r install-mlr3extra, eval=FALSE, message=FALSE, warning=FALSE}
devtools::install_github("mlr-org/mlr3extralearners")
```

## Datasets

The **interflex** package incorporates several datasets, all accessible via `data()`:

```{r ch1-load-data}
library(interflex)
data(interflex)
ls()
```

### Simulation datasets

- **`s1`**: Dichotomous treatment, linear CME. Used in @sec-extensions (fixed effects, multiple arms, predictions, treatment effect differences).
- **`s2`**: Continuous treatment, linear CME. Used in @sec-extensions.
- **`s3`**: Dichotomous treatment, nonlinear (quadratic) CME. Used in @sec-extensions.
- **`s4`**: Dichotomous treatment, nonlinear CME with two-way fixed effects. Used in @sec-extensions (fixed effects demos).
- **`s5`**: Discrete treatment (3 arms: A, B, C), nonlinear CME. Used in @sec-extensions (multiple treatment arms, predictions, treatment effect differences).
- **`s6`–`s9`**: Discrete outcome (binary $Y$) simulation datasets with various DGPs. Used in @sec-discrete-outcome.

### Empirical datasets

- **`app_hma2015`**: Data from @huddy2015expressive. Binary treatment, continuous moderator. Used in @sec-classic (binary treatment demos).
- **`app_vernby2013`**: Data from @vernby2013inclusion. Continuous treatment, continuous moderator. Used in @sec-classic (continuous treatment demos).
- **`app_adiguzel2023`**: Data from @adiguzel2023out. Continuous treatment (change in clinic congestion), continuous moderator (log property value). Used in @sec-classic, @sec-lasso, and @sec-dml (continuous treatment demos).
- **`app_bb2024`**: Data from @beiser2024pocketbook. Binary treatment, discrete moderator (income categories). Used in @sec-extensions (GATE with discrete moderator).
- **`app_et2023`**: Data from @egerod2023extremists. Binary treatment, continuous moderator. Used in @sec-discrete-outcome.



