# Classic: Basics {#sec-classic}

In this chapter, we show how to use **interflex** package to analyze and estimate CME. We begin with visual diagnostics such as subgroup scatterplots, binning plot, and Generalized Additive Models (GAM) plot. We then estimate CME with classic approaches, including linear estimator and kernel estimator; We also show how to use **interflex** package with binning, linear and kernel estimators when users are interested in extension function, such as multiple treatment arms, fixed effect, prediction, and analyzing differences in treatment effects. We illustrates the implementation details using empirical example, with binary treatment settings and continuous treatment settings.

<!-- and 2. outcome-based approach, including IPW estimator, and AIPW estimator. . -->

## Binary treatment

In this example and through out the whole manual, we will use a consistent notation. $Y$ is the outcome variable, the moderator is $X$, the covariates are $Z$.

The example we are using is from @huddy2015expressive, which explores the effect of the expressive model of partisanship. Drawing on four survey studies, the authors argue that partisan identity drives campaign participation and strong emotional responses to ongoing campaign events. The Conditional Marginal Effect (CME) studied in the paper suggests that strongly identified partisans are more likely to exhibit stronger emotional reactions – such as anger when threatened with electoral loss - compared to those with weaker partisan identities. The variables of interest include:

-   Outcome variable: level of anger (continuous scale from 0 to 1)

-   Treatment variable: electoral loss threat or electoral win reassurance (binary yes/no variable)

-   Moderator Variable: strength of partisan identity (continuous scale from 0 to 1)

```{r ch2-init, include=FALSE}
source("init.R")
```

```{r ch2-load-data, message = FALSE, warning = FALSE}
library(interflex)
data(interflex)
ls()
```

```{r ch2-binary-data-setup, message=FALSE, warning=FALSE, cathe = TRUE}
d<-app_hma2015

Y = "totangry" ## outcome: anger
D = "threat" ## treatment: threat
X = "pidentity" ## moderator: partisan identity 
Z <-  c("issuestr2", "knowledge", "educ","male", "age10")

Ylabel = "Anger"
Dlabel = "Threat"
Xlabel = "Partisanship"

datatable(d)
```

------------------------------------------------------------------------

### Diagnostics

#### Raw plots

The first step of the diagnostics is to plot raw data. We supply the function **interflex** with the variable names of the outcome `Y`, the treatment `D`, and the moderator `X`, and covaraites `Z`. You can also supply labels for these variables, such as `Ylabel`, `Dlabel`, and `Xlabel`. If you supply a variable name to the `weights` option, the linear and LOESS fits will be adjusted based on the weights. Note that the correlations between covariates `Z` and `Y` are NOT partialled out.

<!-- We use `main` to add a title to the plot, `cex.main` to adjust its size, and `ylab` to change the y-axis label. -->

```{r ch2-binary-raw-plot, message=FALSE, warning=FALSE, cathe = TRUE, cache=FALSE, fig.width = 6, fig.height = 4.5}

interflex(estimator = "raw",data = d, Y = Y, 
                     D = D, X = X, Z = Z,
                     Ylabel = Ylabel, Dlabel = Dlabel,
                     Xlabel = Xlabel,
          main = "Raw plot", ylab = "CME of Threat on Anger")
```

<!-- A black-white theme is applied when we set `theme.bw = TRUE`. `show.grid = FALSE` can be used to remove grid in the plot. Both options are allowed in **interflex** and **plot**. -->

<!-- ```{r} -->

<!-- interflex(estimator = "raw", data = d, Y = Y,  -->

<!--                      D = D, X = X, Z = Z, -->

<!--                      Ylabel = "Anger", Dlabel = "Threat", -->

<!--                      Xlabel = "Partisanship", -->

<!--           show.grid = FALSE) -->

<!-- ``` -->

#### Binning plots

The second diagnostic tool is the binning plot. It will discretizes the moderator $X$ into several bins and create a dummy variable for each bin and picks an evaluation point within each bin, where users want to estimate the CME of $D$ on $Y$. The regression terms includes interactions between the bin dummies and the treatment indicator, the bin dummies and the moderator $X$ minus the evaluation points, as well as the triple interactions.

The histograms at the bottom of the plot depict the distributions of $X$ across treatment (pink) and control (gray) group. If the CME estimates from the binning estimator sit closely to the estimated CME from the linear estimator, then the Linear Interaction Effect (LIE) assumption likely holds

The `nbins` option sets the number of bins. The default number of bins is 3, and equal-sized bins are created based on the distribution of the moderator.

```{r ch2-binary-bin1-fit, message=FALSE, warning=FALSE, cathe = TRUE, cache=TRUE}
bin.out.1 <- interflex(estimator = "binning", data = d,
                 Y = Y,D = D, X = X, Z = Z,
                 Ylabel = Ylabel, Dlabel = Dlabel,
                 Xlabel = Xlabel,
                 main = "Marginal Effects")
```

```{r ch2-binary-bin1-plot, cache=FALSE, fig.width = 6, fig.height = 4.5}
bin.out.1$figure
```

For fully moderated model, we includes interactions between each covariate $Z$ and the bin dummies and the triple interactions. It can reduce biases when the impact of some covariates to the outcome does not take a linear form of moderator $X$, but sometimes it will also render the estimates unstable when the number of covariates is large.

```{r ch2-binary-bin2-fit, message=FALSE, warning=FALSE, cathe = TRUE, cache=TRUE}
bin.out.2 <- interflex(estimator = "binning", data = d,
                 Y = Y,D = D, X = X, Z = Z,
                 Ylabel = Ylabel, Dlabel = Dlabel,
                 Xlabel = Xlabel, full.moderate = TRUE,
                 main = "Marginal Effects")
```

```{r ch2-binary-bin2-plot, cache=FALSE, fig.width = 6, fig.height = 4.5}
bin.out.2$figure
```

Note that we can customize the cutoff values for the bins, for example, set `cutoffs = c(0.25,0.5,0.75,0.9)` to create five bins: $[minX, 0.25], (0.25, 0.5], (0.5,0.75], (0.75, 0.9]$ and $(0.9,maxX]$ (supplying N numbers will create N+1 bins). Note that the `cutoffs` option will override the **nbins** option if they are incompatible.

```{r ch2-binary-bin3-fit, message=FALSE, warning=FALSE, cathe = TRUE, cache=TRUE}
cutoff<-c(0.25,0.5,0.75,0.9)

bin.out.3 <- interflex(estimator = "binning", data = d,
                 Y = Y,D = D, X = X, Z = Z,
                 Ylabel = Ylabel, Dlabel = Dlabel,
                 Xlabel = Xlabel, nbins = 5,
                 main = "Marginal Effects")
```

```{r ch2-binary-bin3-plot, cache=FALSE, fig.width = 6, fig.height = 4.5}
bin.out.3$figure
```

Next, we use `Xunif = TRUE` to transform the moderator into a uniformly distributed random variable (based on the rank order in values of the orginal moderator) before estimating the CME.

```{r ch2-binary-bin4-fit, message=FALSE, warning=FALSE, cathe = TRUE, cache=TRUE}
bin.out.4 <- interflex(estimator = "binning", data = d,
                 Y = Y,D = D, X = X, Z = Z,
                 Ylabel = Ylabel, Dlabel = Dlabel,
                 Xlabel = Xlabel, Xunif = TRUE,
                 main = "Marginal Effects")
```

```{r ch2-binary-bin4-plot, cache=FALSE, fig.width = 6, fig.height = 4.5}
bin.out.4$figure
```

**interflex** will also automatically report a set of statistics when `estimator = "binning"`, including: (1) the binning estimates and their standard errors and 95% confidence intervals, (2) the percentage of observations within each bin, (3) the L-kurtosis of the moderator, and (4) a Wald test to formally test if we can reject the linear multiplicative interaction model by comparing it with a more flexible model of multiple bins.

The CME plot is stored in `object$figure` while the estimates and standard errors are stored in `object$est.bin` (binning). The tests results are stored in `object$tests` (binning).

```{r ch2-binary-bin-wald-test, message=FALSE, warning=FALSE, cathe = TRUE}
print(bin.out.1$tests$p.wald)
```

`plot` allows users to adjust graphic options without re-estimating the model. The first entry must be a **interflex** object. Note that we use `bin.labs = FALSE` to hide the label on the top of each bin and `Xdistr = "none` to remove the distribution of the moderator (not recommended), or set `Xdistr = "density` to present the distribution of the moderator with a density plot using option. We use `cex.axis` and `cex.lab` to adjust the font sizes of axis numbers and labels. We address all the plotting options in chapter @sec-plotting.

```{r ch2-binary-bin-replot, message=FALSE, warning=FALSE, cathe = TRUE, cache=FALSE, fig.width = 6, fig.height = 4.5}

plot(bin.out.1, Xdistr = "density", bin.labs = FALSE)
```

------------------------------------------------------------------------


### Linear estimator

We first discuss classic approaches to estimating and visualizing the CME. We begin with the linear interaction model, which remains widely used in social science research.


**Estimation.** If we want to deploy the linear estimator, we can set `estimator = "linear"`. This method estimates CME with multiplicative interaction model. When the option `full.moderate` is set to `TRUE`, interaction terms between the moderator $X$ and each covariate $Z$ will enter the model, which is the case in Blackwell and Olson (2020). This extension in some cases can reduce the bias of the original model.

**Uncertainty estimation.** Depending on `vartype`, the estimator computes standard errors and confidence intervals in three ways `simu`, `delta`, and `bootstrap`. The option `vartype = "simu"` draws from a multivariate normal distribution multiple times, `vartype = "delta"` uses a matrix derivatives to approximate the variance.

We report the uniform/simultaneous confidence interval along with pointwise confidence interval. Pointwise CI is estimated by default, and is depicted by gray ribbon; Uniform CI is estimated when setting `vartype = "bootstrap"` and `vartype = "delta"`, as depicted by the dashed lines.

Note that the variance-covariance matrix of coefficients used by the delta method or the simulation method by default is the robust standard error matrix estimated from the regression. Users can also choose `homoscedastic` or `cluster` standard error by specifying the option `vcov.type`. There are four options for the choice of the `vcov` estimator: `vcov.type = "homoscedastic"`, `"robust"`, `"cluster"`, and `"pcse"`. The default option is `"robust"`.

```{r ch2-binary-lin-fit, message=FALSE, warning=FALSE, cathe = TRUE, cache=TRUE}
linear.out.1 <- interflex(estimator = "linear", data = d,
                 Y = Y,D = D, X = X, Z = Z,
                 Ylabel = Ylabel, Dlabel = Dlabel,
                 Xlabel = Xlabel, vartype = "bootstrap",
                 main = "Linear model")

linear.out.2 <- interflex(estimator = "linear", data = d,
                 Y = Y,D = D, X = X, Z = Z,
                 Ylabel = Ylabel, Dlabel = Dlabel,
                 Xlabel = Xlabel,
                 vartype = "bootstrap", full.moderate = TRUE,
                 main = "Fully Moderated Linear model")
```

```{r ch2-binary-lin-plot, cache=FALSE, fig.width = 9, fig.height = 5}
ggarrange(linear.out.1$figure, linear.out.2$figure)
```

::: {.callout-tip}
## Parallel computing

Parallel computing will speed up both cross-validation and uncertainty estimation significantly. We recommend that users manually set the number of cores using the `cores` option. If this is not supplied or is `NULL`, **interflex** will automatically select the smaller of 8 and the number of usable system cores minus 2, to prevent excessive use of system resources.
:::

**Results.** The CME plot is stored in `out$figure` while the estimates and standard errors are stored in `out$est.lin`

```{r ch2-binary-lin-results, message=FALSE, warning=FALSE, cathe = TRUE}
head(linear.out.1$est.lin$`1`)
```

### Uniform confidence intervals

By default, **interflex** reports both pointwise and uniform confidence intervals (CIs) when `vartype = "bootstrap"`. The pointwise CI covers the true CME at each individual value of $X$ with 95% probability, but it does *not* guarantee simultaneous coverage across all values of $X$. The uniform CI, by contrast, provides a confidence band that covers the *entire* CME curve simultaneously with 95% probability. This is analogous to the distinction between individual and family-wise error rates in multiple testing.

Uniform CIs are wider than pointwise CIs. They are important when making statements about the shape of the CME curve as a whole — for example, "the treatment effect is positive for all values of $X > 0$" requires the uniform band to be above zero for the entire region of $X > 0$, not just each pointwise interval. In the plots, pointwise CIs appear as shaded ribbons and uniform CIs appear as dashed lines.

::: {.callout-note}
## How many bootstrap samples?

The bootstrapped uniform CI works by finding the narrowest band that achieves simultaneous 95% coverage across all $k$ evaluation points. With $N$ bootstrap draws, the smallest resolvable tail probability is approximately $1/N$, and the algorithm improves on the Bonferroni correction only when $N > 2k/\alpha$. For the default $\alpha = 0.05$:

| Evaluation points ($k$) | Minimum `nboots` |
|---|---|
| 10 | 400 |
| 20 | 800 |
| 50 (default) | 2,000 |
| 100 | 4,000 |

With the default `nboots = 200`, the algorithm falls back to the conservative Bonferroni correction for most cases. We recommend setting `nboots = 2000` or higher when uniform CIs are of interest.
:::

Here we demonstrate the linear estimator with `vartype = "bootstrap"` and `nboots = 2000` to obtain proper uniform CIs:

```{r ch2-uniform-ci-fit, cache=TRUE, message=FALSE, warning=FALSE}
linear.uni <- interflex(estimator = "linear", data = d,
                 Y = Y, D = D, X = X, Z = Z,
                 Ylabel = Ylabel, Dlabel = Dlabel,
                 Xlabel = Xlabel, vartype = "bootstrap",
                 nboots = 2000, main = "With Uniform CI (nboots = 2000)")
```

```{r ch2-uniform-ci-plot, cache=FALSE, fig.width = 6, fig.height = 4.5}
linear.uni$figure
```

To turn off uniform CIs (e.g., to reduce visual clutter or when only pointwise inference is needed), set `show.uniform.CI = FALSE` in either the estimation or plotting call:

```{r ch2-uniform-ci-off-fit, cache=TRUE, message=FALSE, warning=FALSE}
linear.nouni <- interflex(estimator = "linear", data = d,
                 Y = Y, D = D, X = X, Z = Z,
                 Ylabel = Ylabel, Dlabel = Dlabel,
                 Xlabel = Xlabel, vartype = "bootstrap",
                 nboots = 2000,
                 show.uniform.CI = FALSE, main = "Without Uniform CI")
```

```{r ch2-uniform-ci-off-plot, cache=FALSE, fig.width = 6, fig.height = 4.5}
linear.nouni$figure
```

You can also hide uniform CIs at the plotting stage without re-estimating:

```r
plot(linear.uni, show.uniform.CI = FALSE)
```

### Kernel estimator

To relax functional form assumptions, we then explore the kernel estimator proposed by @hainmueller2019much.

**Estimation.** Similar to other estimator, to deploy the kernel estimator, we can set `estimator = "kernel"`. We also provide `full.moderate` option, allowing moderator $X$ and each covariate $Z$ to fully interact.

**Uncertainty estimation.** Kernel estimator computes standard errors in two ways: `delta` and `bootstrap`. Uniform CI is estimated when setting `vartype = "bootstrap"`.

When the standard errors are produced bootstrapping, we can adjust the number of bootstrap iterations using the `nboots` option, with the default being 200 (set `nboots = 2000` or higher for proper uniform CIs; see the note above). The grid option can either be an integer, which specifies the number of bandwidths to be cross-validated, or a vector of candidate bandwidths -- the default is `grid = 20`. We can also specify a clustered group structure by using the `cl` option, in which case a block bootstrap procedure will be performed.


The left panel uses the default settings for the kernel estimator; the right panel use fully interacted model, with `nboots = 400`, and we set the estimator to choose bandwidths from `c(0,1,2,3)`

```{r ch2-binary-kernel-fit, message=FALSE, warning=FALSE, cathe = TRUE, cache=TRUE}
kernel.out.1 <- interflex(estimator = "kernel",
                          data = d,
                 Y = Y,D = D, X = X, Z = Z,
                 Ylabel = Ylabel, Dlabel = Dlabel,
                 Xlabel = Xlabel,
                 vartype = "bootstrap", bw = 0.916,
                 main = "Kernel model")


kernel.out.2 <- interflex(estimator = "kernel",
                          data = d,
                 Y = Y,D = D, X = X, Z = Z,
                 Ylabel = Ylabel, Dlabel = Dlabel,
                 Xlabel = Xlabel,
                 vartype = "bootstrap", grid = c(0,1,2,3), full.moderate = TRUE,
                 main = "Fully Moderated Kernel model")
```

```{r ch2-binary-kernel-plot, cache=FALSE, fig.width = 9, fig.height = 5}
ggarrange(kernel.out.1$figure, kernel.out.2$figure)
```

**Bandwidth selection.** We use adaptive bandwidth selection in **interflex**. This means that the bandwidth will be smaller in regions where there are more data and larger in regions where there are few data points, based on the moderator $X$. When adaptive bandwidth is being used, `bw` refers to the bandwidth applied in the region with the highest density of observations.

Even with an optimized algorithm, the bootstrap procedure for large datasets can be slow. We incorporate parallel computing, `parallel=TRUE`, to speed it up. You can choose the number of cores to be used for parallel computing; otherwise the program will detect the number of logical cores in your computer and use as many cores as possible (warning: this will significantly slow down your computer!).

If you specify a bandwidth manually, for example, by setting `bw = 1`, the cross-validation step will be skipped.

```{r ch2-binary-kernel-bandwidth, message=FALSE, warning=FALSE, cathe = TRUE, cache=TRUE}
cat("1. Runtime with parallel computing for Bandwidth selection:\n")
runtime_parallel <- system.time({
  kernel.out.parallel <- interflex(estimator = "kernel", data = d, 
                                   Y = Y, D = D, X = X, Z = Z,
                                   Ylabel = Ylabel, Dlabel = Dlabel,
                                   Xlabel = Xlabel,
                                   vartype = "bootstrap", parallel = TRUE,
                                   main = "Linear model")
})
cat("User Time: ", runtime_parallel[1], " seconds\n")
cat("System Time: ", runtime_parallel[2], " seconds\n")
cat("Elapsed Time: ", runtime_parallel[3], " seconds\n\n")

cat("2. Runtime with user-specified Bandwidth selection:\n")
runtime_user_bandwidth <- system.time({
  kernel.out.user_bandwidth <- interflex(estimator = "kernel", data = d, 
                                         Y = Y, D = D, X = X, Z = Z,
                                         Ylabel = Ylabel, Dlabel = Dlabel,
                                         Xlabel = Xlabel,
                                         bw = 0.9,
                                         main = "Linear model")
})
cat("User Time: ", runtime_user_bandwidth[1], " seconds\n")
cat("System Time: ", runtime_user_bandwidth[2], " seconds\n")
cat("Elapsed Time: ", runtime_user_bandwidth[3], " seconds\n")
```

**Results.** The estimation results, including pointwise and uniform confidence intervals, are included in `est.kernel`.

```{r ch2-binary-kernel-results, message=FALSE, warning=FALSE, cathe = TRUE}
head(kernel.out.1$est.kernel$`1`)
```

The results of cross-validation are saved in `CV.output`.

```{r ch2-binary-kernel-cv, message=FALSE, warning=FALSE, cathe = TRUE}
head(kernel.out.1$CV.output)
```

------------------------------------------------------------------------

## Continuous treatment

@adiguzel2023out examines the impact of incumbent's public health provision on electoral outcomes. The authors use a reform in Turkey that significantly altered the geographic distribution of health clinics. The authors find that reduced congestion levels significantly increased the AKP's vote share, especially in poorer communities which exhibited a more substantial response to the improvements in health care access. The variables of interest are:

-   Outcome: Change in the vote share for the incumbent AKP party (continuous, $\in [-33.9, 39.5]$)
-   Treatment: Change in congestion (continuous, $\in [-28.8, 20.8]$); congestion is measured by the number of patients per doctor at the nearest clinic, with lower congestion indicating improved service quality
-   Moderator: Logarithm of median property value (continuous, $\in[-0.2, 7.7]$)

```{r ch2-con-data-setup, message=FALSE, warning=FALSE, cathe = TRUE}
data(app_adiguzel2023)
d<-app_adiguzel2023
d<-remove_labels(d)
d<-as.data.frame(d)
main<-"Adiguzel2023a"
D="Dwalk_3" 
Y="Dakp_3" 
X="lograyic09"

Z=c('Duniversity_3', 'Dodr_3', 'Dydr_3', 'Dpopulation_3', 'Dhospital_pri_3', 'Dhospital_pub_3')

Dlabel<-"Dwalk"
Ylabel<-"Dakp"
Xlabel<-"Property values (2009)"

#datatable(d)
```

------------------------------------------------------------------------

### Diagnostics

#### Raw plots

Similar to the binary treatment settings, we can plot the subgroup scatter plot as first step of diagnostics:

```{r ch2-con-raw1-plot, message=FALSE, warning=FALSE, cathe = TRUE, cache=FALSE, fig.width = 6, fig.height = 4.5}
interflex(estimator = "raw",data = d, Y = Y, 
                     D = D, X = X, Z = Z,
                     Ylabel = Ylabel, Dlabel = Dlabel,
                     Xlabel = Xlabel,
          main = "Raw plot", ylab = "CME")
```

Instead of grouping by treatment status, the scatter plot carves $X$ into 3 quantile‐based bins by default. You can supply your own number of bins by setting the `cutoff =`. Supplying $N$ numbers will create $N+1$ bins.

```{r ch2-con-raw2-plot, message=FALSE, warning=FALSE, cathe = TRUE, cache=FALSE, fig.width = 6, fig.height = 4.5}
cut_bins = c(3, 4, 5)
interflex(estimator = "raw",data = d, Y = Y, 
                     D = D, X = X, Z = Z,
                     Ylabel = Ylabel, Dlabel = Dlabel,
                     Xlabel = Xlabel, cutoff = cut_bins,
          main = "Raw plot", ylab = "CME")
```

#### Binning plots

Similar to the binary treatment case, we can also use binning plot for diagnosis.

```{r ch2-con-bin-fit, message=FALSE, warning=FALSE, cathe = TRUE, cache=TRUE}
bin.out.1<-interflex(estimator = "binning",data = d, Y = Y,
                     D = D, X = X, Z = Z,
                     Ylabel = Ylabel, Dlabel = Dlabel,
                     Xlabel = Xlabel, cutoff = cut_bins,
          main = "Binning plot", ylab = "CME")
```

```{r ch2-con-bin-plot, cache=FALSE, fig.width = 6, fig.height = 4.5}
bin.out.1$figure
```

Because the treatment group is not binary, the histograms at the bottom of the plot simply depict the distribution of $X$, not differentiating by treatment status.

#### GAM plots

For the continuous treatment case, we can draw a Generalized Additive Model (GAM) plot.

```{r ch2-con-gam1-plot, message=FALSE, warning=FALSE, cathe = TRUE, eval=FALSE, cache=FALSE, fig.width = 6, fig.height = 4.5}
interflex(estimator = "gam",data = d, Y = Y, 
                     D = D, X = X, Z = Z,
                     Ylabel = Ylabel, Dlabel = Dlabel,
                     Xlabel = Xlabel, show.uniform.CI = FALSE,
          main = "GAM plot", ylab = "CME")
```

You can manipulate the angle of each of the four plots by setting `angle` with your own value of theta, which change the azimuth (horizontal rotation) angle in degrees. The default angle is `c(30, 100,-30,-120).`

```{r ch2-con-gam2-plot, message=FALSE, warning=FALSE, cathe = TRUE, cache=FALSE, fig.width = 6, fig.height = 4.5}
interflex(estimator = "gam",data = d, Y = Y, 
                     D = D, X = X, Z = Z,
                     Ylabel = Ylabel, Dlabel = Dlabel,
                     Xlabel = Xlabel, angle = c(40, 50, 100, 120),
          main = "Raw plot", ylab = "CME")
```

------------------------------------------------------------------------

### Linear estimator

We illustrate the code implementation using an empirical example with a continuous treatment. Note that the coding approach is essentially the same for both continuous and binary treatments.

```{r ch2-con-lin-fit, message=FALSE, warning=FALSE, cathe = TRUE, cache=TRUE}
linear.out.1 <- interflex(estimator = "linear", data = d,
                 Y = Y,D = D, X = X, Z = Z,
                 Ylabel = Ylabel, Dlabel = Dlabel,
                 Xlabel = Xlabel,
                 main = "Linear model")

linear.out.2 <- interflex(estimator = "linear", data = d,
                 Y = Y,D = D, X = X, Z = Z,
                 Ylabel = Ylabel, Dlabel = Dlabel,
                 Xlabel = Xlabel, full.moderate = TRUE,
                 main = "Fully Moderated Linear model")
```

```{r ch2-con-lin-plot, cache=FALSE, fig.width = 9, fig.height = 5}
ggarrange(linear.out.1$figure, linear.out.2$figure)
```

### Kernel estimator

We can deploy kernel estimator with the same syntax.

```{r ch2-con-kernel-fit, message=FALSE, warning=FALSE, cathe = TRUE, cache=TRUE}
kernel.out.1 <- interflex(estimator = "kernel",
                          data = d,
                 Y = Y,D = D, X = X, Z = Z,
                 Ylabel = Ylabel, Dlabel = Dlabel,
                 Xlabel = Xlabel,
                 vartype = "bootstrap",
                 main = "Kernel model")

kernel.out.2 <- interflex(estimator = "kernel",
                          data = d,
                 Y = Y,D = D, X = X, Z = Z,
                 Ylabel = Ylabel, Dlabel = Dlabel,
                 Xlabel = Xlabel,
                 vartype = "bootstrap", grid = c(0,1,2,3), full.moderate = TRUE,
                 main = "Fully Moderated Kernel model")
```

```{r ch2-con-kernel-plot, cache=FALSE, fig.width = 9, fig.height = 5}
ggarrange(kernel.out.1$figure, kernel.out.2$figure)
```

