# Plot Options {#sec-plotting}

In previous chapter, we have explored most of the plotting options in the **interflex** package. This chapter serves as a comprehensive guide in plotting CME, using data from @huddy2015expressive and simulation data incorporated in the package.

## Load Data

In @huddy2015expressive, the outcome variable is level of anger, treatment variable is electoral loss threat or electoral win reassurance, and moderator $X$ is strength of partisan identity. The data is included in the **interflex** package. We also specify the treatment, outcome, moderator, and controls (covariates).

```{r include=FALSE}
source("init.R")
```

```{r ch7-data-setup}
data(interflex)
ls()

d<-app_hma2015

## Define variables 
Y = "totangry" ## outcome: anger
D = "threat" ## treatment: threat
X = "pidentity" ## moderator: partisan identity 
Z <-  c("issuestr2","knowledge", "educ","male", "age10")

Ylabel = "Anger"
Dlabel = "Threat"
Xlabel = "Partisanship"
```

## The plot function

Users have two equally flexible ways to customise figures produced by interflex:

When setting the option `show.all` to TRUE in the command plot, the program will return a list in which every element is the plot of one treated group. As all of them are ggplot2 objects, users can then edit those graphs in their ways.

1.  During estimation: pass any of the plotting-related arguments into `interflex()`, then retrieve the figure.

```{r fig.width = 6, fig.height = 4.5}
linear.out.1 <- interflex(estimator = "linear", data = d, 
                 Y = Y,D = D, X = X, Z = Z,
                 Ylabel = Ylabel, Dlabel = Dlabel,
                 Xlabel = Xlabel, vartype = "bootstrap",
                 main = "Linear model")
linear.out.1$figure
```

2.  After estimation: call `plot(result, …)` and supply or overwrite the same arguments.

```{r fig.width = 6, fig.height = 4.5}
linear.out.1 <- interflex(estimator = "linear", data = d, 
                 Y = Y,D = D, X = X, Z = Z,
                 Ylabel = Ylabel, Dlabel = Dlabel,
                 Xlabel = Xlabel, vartype = "bootstrap",
                 main = "Linear model")
plot(linear.out.1)
```

All the options introduced in this chapter are available for both ways.

### Estimates to plot

`order`: For multi-treatment plots: character (discrete treatment) or numeric (continuous) vector that fixes the display order of treatment arms / reference values.

`pool`: If you supply a list of **interflex** objects (x), set `pool = TRUE` to draw a single pooled plot across objects.

For instance, with simulation data **s5**, by setting the option `order` and `subtitles`, users can also specify the order of plots they prefer and the subtitles they want:

```{r mta1c, message = FALSE, fig.height=5, fig.width=8, cache = TRUE,results = "hide", fig.width = 6, fig.height = 4.5}

s5.binning <- interflex(Y = "Y", D = "D", X = "X", Z = c("Z1","Z2"), 
                        data = s5, 
                        estimator = "binning", vartype = "bootstrap")

plot(s5.binning, order=c("C", "B"), 
     subtitles = c("Group C","Group B"))
```

By setting `pool = TRUE`, we can combine these plots. Users can also specify the colors they like:

```{r mta1d, message = FALSE, fig.height=5, fig.width=6, cache = TRUE}
plot(s5.binning, order=c("C", "B"), 
     subtitles = c("Control Group","Group C","Group B"), 
     pool = TRUE, color = c("Salmon","#7FC97F", "#BEAED4"))
```


`diff.values`: **interflex** allows users to compare treatment effects at three specific values of the moderator using linear or kernel model. In order to do so, users need to pass three values into `diff.values` when applying *linear* or *kernel* estimator.

`CI`: TRUE / FALSE (or the numeric level) to draw pointwise confidence bands. The default is estimator-dependent (e.g. the kernel estimator uses whatever was saved in the object).

Using the linear estimator with simulation data **s5** as an example:

```{r ch7-mta-linear-fit, cache=TRUE, message=FALSE, warning=FALSE}
s5.linear <-interflex(Y = "Y", D = "D", X = "X", Z = c("Z1","Z2"),
                      data = s5,
                      estimator = "linear")
```

```{r ch7-mta-linear-plot, cache=FALSE}
plot(s5.linear,
     diff.values = c (-2,0,2),
     CI = FALSE)
```

### Moderator's distribution

`Xdistr = "histogram"`: Adds a stacked treatment/control histogram, a shaded marginal density, or nothing beneath the effect curve(s). Options are `"histogram"`, `"density"`, and `"none"`.

`hist.color`/`density.color`/`hist.color.alpha`/`density.color.alpha`: adjust color and Transparency of the histogram bars / density ribbons.

Using the linear estimation plot with Huddy2015 as an example:

```{r fig.width = 6, fig.height = 4.5}
linear.out.1 <- interflex(estimator = "linear", data = d, 
                 Y = Y,D = D, X = X, Z = Z,
                 Ylabel = Ylabel, Dlabel = Dlabel,
                 Xlabel = Xlabel, vartype = "bootstrap",
                 main = "Linear model")

plot(linear.out.1, 
     Xdistr = "density",
     density.color = "blue")
```

### Titles, annotations & facet subtitles

`main`: Main title (top-center).

`subtitles`: Vector of custom facet subtitles (length = number of non-baseline treatment arms or reference points).

`show.subtitles`: Logical; force subtitles on/off irrespective of subtitles or panel count.

`interval`: Numeric(s); draws light-blue vertical reference line(s) at those moderator values.

`legend.title`: Title for the legend generated when `pool = TRUE`.

Using the linear estimation plot with Huddy2015 as an example:

```{r ch7-mta-linear-subtitle-plot, cache=FALSE}
plot(s5.linear,
     interval = c(0),
     show.subtitles = TRUE,
     main = "Linear estimator",
     subtitles = c("Group B","Group C"))
```

### Axis labels & limits

`xlab`/`ylab`: Full axis labels (default glue: "Moderator: Xlabel", "Marginal Effect of Dlabel on Ylabel").

`xlim`/`ylim`: Numeric length-2 vectors to modify the x-axis and y-axis range.

Using the linear estimation plot with Huddy2015, and compare it against the plot without any modification

```{r fig.width = 8, fig.height = 6}

plot1<-linear.out.1$figure

plot2<-plot(linear.out.1, xlab = "moderator", ylab = "CME",
     xlim = c(0.25,1))

ggarrange(plot1, plot2)

```

### Theme and layout

`theme.bw`: TRUE adds theme_bw() and a light grey zero-line; FALSE uses the default ggplot background and a white zero-line.

`show.grid`: Toggles panel major/minor grid lines.

`line.color`/`line.size`: Colour and thickness of the treatment-effect curve(s).

`CI.color`/`CI.color.alpha`: Fill colour + alpha for the pointwise CI ribbon.

`bin.labs`: If you used the binning estimator, prints "L/M/H …" labels underneath each bin centre.

`cex.main (title)`/`cex.sub (subtitle)`/`cex.lab (axis titles)`/`cex.axis (tick labels)`: font size options

`ncols`: Number of columns when multiple panels are arranged. Defaults to the number of panels.

Using the linear estimation plot with Huddy2015 as an example:

```{r fig.width = 6, fig.height = 4.5}
plot(linear.out.1, theme.bw = FALSE,
     show.grid = FALSE,
     CI.color = "red")
```
