# Double Machine Learning {#sec-dml}

This chapter gives a brief overview of estimating CMEs using the **interflex** pacakge, focusing on the Double/debiased Machine Learning (DML) estimator. This chapter walks through using DML estimators for different empirical setup with real-world data, including binary treatment and continuous treatment.

## Overview of Methodology

When treatment $D$, moderator $X$ and covariates$Z$ are all discrete, we can estimate the CME nonparametrically. However, when one of them is continuous or high-dimensional, CME becomes difficult to estimate without imposing functional form assumptions. These assumptions can be too restrictive and easily violated in practice.

Here, we introduce the double/debiased machine learning (DML) estimator, to flexibly estimate CME under mild conditions. The primary advantage of DML lies in its ability to use machine learning method to estimate nuisance parameters $\eta_0$ (parameters that are not of direct interest to researchers), in nonparametric, data-driven way, while providing valid inference for the target quantity, in this case, the CME $\theta_0$.

The DML framework operates through the following steps:

**1. Estimate nuisance functions**

-   Outcome model: Estimate the nuisance function that captures relationship between outcome $Y$ covariate $V = (X, Z)$, using ML methods.
-   Treatment model: Estimate the nuisance function that captures relationship between treatment $D$ covariate $V = (X, Z)$.

**2. Construct signals** The key of DML estimator is the construction of orthogonal score to meet the Neyman orthogonality condition. With binary treatment, the signal is constructed through AIPW correction; With continuous treatment, the signal is constructed via "partialling out".

**3. Dimension reduction**

The constructed orthogonal signal is projected onto a functional space in $X$, which, in the **interflex** package, is either a B-spline or kernel regression, to recover CME.

------------------------------------------------------------------------

```{r include=FALSE}
source("init.R")
```

<!-- Load the simulated toy datasets, `s6`, `s7`, and `s9`, for this tutorial. `s6` is a case of a binary treatment indicator with discrete outcomes. `s7` is a case of a continuous treatment indicator with discrete outcomes. `s9` is a case of a discrete treatment indicator (3 groups) with discrete outcomes. -->

<!-- ```{r} -->

<!-- data(interflex) -->

<!-- ls() -->

<!-- ``` -->

When setting the option `estimator` to `DML`, the function will conduct the two-stage DML estimator to compute the marginal effects. In the following command, the outcome variable is $Y$, the moderator is $X$, the covariates are $Z$.

## Binary treatment

In the binary treatment setting, we are using the @huddy2015expressive example, with outcome variable being level of anger, treatment variable being electoral loss threat or electoral win reassurance, and moderator $X$ is strength of partisan identity.

```{r app1_data, message=FALSE, warning=FALSE, cathe = TRUE}
d<-app_hma2015

Y = "totangry" ## outcome: anger
D = "threat" ## treatment: threat
X = "pidentity" ## moderator: partisan identity 
Z <-  c("issuestr2", "knowledge", "educ","male", "age10")

Ylabel = "Anger"
Dlabel = "Threat"
Xlabel = "Partisanship"
```

### DML estimator

**Estimation.** To implement the DML estimator, we need to set the option `estimator='DML'` and specify two machine learning models: one for the outcome model (via the option `model.y`) and one for the treatment model (via the option `model.t`). `treat.type` set the data type of treatment variable, it can be either `continuous` or `discrete`, the default is `discrete`.

Here, we use a neural network model (nn) for the outcome model and a random forest model (rf) for the treatment model. The gray area depicts the pointwise interval, while the dashed line marks the uniform/joint confidence interval.

```{r app1_nn_rf, message=FALSE, warning=FALSE, cathe = TRUE, fig.width = 6, fig.height = 4.5}
binary.dml.nn.rf <- interflex(estimator="dml", data = d,
                              Y = Y, D = D, X = X, Z = Z,
                              Ylabel = Ylabel, Dlabel = Dlabel,
                              Xlabel = Xlabel,
                              model.y = "nn",model.t = "rf")
plot(binary.dml.nn.rf, theme.bw = TRUE)
```

**Results.** The estimated treatment effects of $D$ on $Y$ across the support of $X$ are saved in:

```{r}
## printing the first 10 rows
lapply(binary.dml.nn.rf$est.dml, head, 10)
```

### First Stage DML Model Selections

The `DML` estimator supports five different machine learning algorithms: `linear` (Linear model), `regularization` (Lasso or Ridge model),`rf` (Random Forest), `nn` (Neural Network), and `hgb` (Histogram-based Gradient Boosting). Users can change the `model.y` and `model.t` option to calculate the nuisance parameter using a different ML method.

```{r app1_firstage, message=FALSE, warning=FALSE, cathe = TRUE}

## linear + linear
binary.dml.ln.ln <- interflex(estimator="DML", data = d,
                              Y = Y, D = D, X = X, Z = Z,
                              Ylabel = Ylabel, Dlabel = Dlabel,
                              Xlabel = Xlabel, theme.bw = TRUE,
                              model.y = "linear",model.t = "linear",
                              main = "Linear Model")

# # lass + lasso
# binary.dml.ls.ls <- interflex(estimator="DML", data = d,
#                               Y = Y, D = D, X = X, Z = Z,
#                               Ylabel = Ylabel, Dlabel = Dlabel,
#                               Xlabel = Xlabel, theme.bw = TRUE,
#                               model.y = "regularization",
#                               model.t = "regularization",
#                               main = "Lasso Model")


## neural network + neural network
binary.dml.nn.nn <- interflex(estimator="DML", data = d,
                              Y = Y, D = D, X = X, Z = Z,
                              Ylabel = Ylabel, Dlabel = Dlabel,
                              Xlabel = Xlabel, theme.bw = TRUE,
                              model.y = "nn",model.t = "nn",
                              main = "Neural Network")


## random forest + random forest
binary.dml.rf.rf <- interflex(estimator="DML", data = d,
                              Y = Y, D = D, X = X, Z = Z,
                              Ylabel = Ylabel, Dlabel = Dlabel,
                              Xlabel = Xlabel, theme.bw = TRUE,
                              model.y = "rf",model.t = "rf",
                              main = "Random Forest")

## hist gradient boosting + hist gradient boosting
binary.dml.hgb.hgb <- interflex(estimator="DML",data = d,
                              Y = Y, D = D, X = X, Z = Z,
                              Ylabel = Ylabel, Dlabel = Dlabel,
                              Xlabel = Xlabel, theme.bw = TRUE,
                              model.y = "hgb",model.t = "hgb",
                              main = "Hist Gradient Boosting")
```

```{r, fig.width=9,fig.height=5}
ggarrange(binary.dml.ln.ln$figure,
          #binary.dml.ls.ls$figure,
          binary.dml.nn.nn$figure, 
          binary.dml.rf.rf$figure,
          binary.dml.hgb.hgb$figure,
          ncol = 2)
```

We find that different ML models give very similar results, which is expected since all methods can accurately partial out their impacts on the outcome and the treatment.

### Specifying Model Parameters

Users can customize parameters for the machine learning model in both stages by specifying the options `param.y` and `param.t`. For the linear method, users can refer to the [LinearRgression](https://scikit-learn.org/stable/modules/generated/sklearn.linear_model.LinearRegression.html) and [LogisticRegression](https://scikit-learn.org/stable/modules/generated/sklearn.linear_model.LogisticRegression.html) if outcome is discrete. For the lasso method, please refere to [LogisticRegression](https://scikit-learn.org/stable/modules/generated/sklearn.linear_model.LogisticRegression.html) with default regularization. For the whole list of parameters that are applicable to the neural network model, users can refer to the [MLPRegressor](https://scikit-learn.org/stable/modules/generated/sklearn.neural_network.MLPRegressor.html) or the [MLPClassifier](https://scikit-learn.org/stable/modules/generated/sklearn.neural_network.MLPClassifier.html). For the random forest method, users can refer to the [RandomForestRegressor](https://scikit-learn.org/stable/modules/generated/sklearn.ensemble.RandomForestRegressor.html) or the [RandomForestClassifier](https://scikit-learn.org/stable/modules/generated/sklearn.ensemble.RandomForestClassifier.html). For the hist gradient boosting method, users can refer to the [HistGradientBoostingRegressor](https://scikit-learn.org/stable/modules/generated/sklearn.ensemble.HistGradientBoostingRegressor.html) or the [HistGradientBoostingClassifier](https://scikit-learn.org/stable/modules/generated/sklearn.ensemble.HistGradientBoostingClassifier.html).

Here we use Huddy (2015) to explore how selecting different DML parameters affects performance. We can directly specify the size of the neural networks, the activation function, learning rate, and solver using the list `param_nn` and feed them to the DML estimator. We compare it with the DML estimator using nn as first stage model, but without adjusting parameters.

```{r model_parameter, message=FALSE, warning=FALSE, cathe = TRUE, fig.width=9,fig.height=5}
param_nn <- list(
  hidden_layer_sizes = list(10L,20L,10L),
  activation = c("tanh"),
  solver = c("sgd"),
  alpha = c(0.05),
  learning_rate = c("adaptive")
)

binary.DML.nn.linear.para <- interflex(estimator="DML", data = d,
                              Y = Y, D = D, X = X, Z = Z,
                              Ylabel = Ylabel, Dlabel = Dlabel,
                              Xlabel = Xlabel, theme.bw = TRUE,
                              model.y = "nn", param.y = param_nn,
                              model.t = "nn", param.t = param_nn,
                              main = "Specified parameters")

ggarrange(binary.dml.nn.nn$figure,
          binary.DML.nn.linear.para$figure)
```

<!-- ```{r, fig.width=6, fig.height=6} -->

<!-- n <- nrow(s7) -->

<!-- d2 <- s7$D -->

<!-- x_values <- sort(s7$X)  -->

<!-- d2_value <- median(d2)  # Example value for d2 -->

<!-- marginal_effect <- function(D_value, X_value) { -->

<!--   link_value <- -1 + D_value + X_value + D_value * X_value -->

<!--   prob_value <- exp(link_value) / (1 + exp(link_value)) -->

<!--   marginal_effect_value <- prob_value * (1 - prob_value) * (1 + X_value) -->

<!--   return(marginal_effect_value) -->

<!-- } -->

<!-- # Applying the function to the range of x values to calculate true ME -->

<!-- ME <- sapply(x_values, function(x_val) marginal_effect(d2_value, x_val)) -->

<!-- p.s7.nn.linear <- plot(s7.DML.nn.linear.para, show.all = TRUE, ylim = c(-0.5,0.5))[[1]] -->

<!-- p.s7.nn.linear<-p.s7.nn.linear + ylim(-0.5,0.5) +  -->

<!--   geom_line(aes(x=x_values,y=ME),color='red')+ -->

<!--   ggtitle("Neural Network with Specified Parameters")+ -->

<!--   theme(plot.title = element_text(size=10)) -->

<!-- ``` -->

### Cross-Validation

Users can also specify the grids of parameters using the options `param.grid.y` and `param.grid.t`. The DML estimator will then search through all sequences of parameter settings and pick the most suitable one using cross-validation. We need to set the options `CV` to `TRUE` to invoke the cross-validation. As an example, we specify the grids used for the lasso/ridge regression, neural network, random forest, and hist gradient boosting, then we compare their performances on @huddy2015expressive.

```{r binary_search,  message=FALSE, warning=FALSE, cathe = TRUE}
param_grid_nn <- list(
  hidden_layer_sizes = list(list(10L,20L,10L), list(20L)),
  activation = c("tanh","relu"),
  solver = c("sgd","adam"),
  alpha = c(0.0001,0.05),
  learning_rate = c("constant","adaptive")
)

param_grid_rf <- list(
  max_depth = c(2L,3L,5L,7L),
  n_estimators = c(10L, 30L, 50L, 100L, 200L, 400L, 600L, 800L, 1000L),
  max_features = c(1L,2L)
)

param_grid_hgb <- list(
  max_leaf_nodes = c(5L,10L,20L,30L),
  min_samples_leaf = c(5L,10L,20L),
  l2_regularization = c(0,0.1,0.5),
  max_features = c(0.5,1)
)


# nn (cv) + rf (cv) + linear
binary.cvnn.cvrf <- interflex(estimator="DML", data = d,
                              Y = Y, D = D, X = X, Z = Z,
                              Ylabel = Ylabel, Dlabel = Dlabel,
                              Xlabel = Xlabel, theme.bw = TRUE,CV = TRUE,
                              model.y = "nn",param.grid.y = param_grid_nn,
                              model.t = "rf",param.grid.t = param_grid_rf)

# nn (cv) + hgb (cv) + regularization (cv)
binary.cvnn.cvhgb <- interflex(estimator="DML", data = d,
                              Y = Y, D = D, X = X, Z = Z,
                              Ylabel = Ylabel, Dlabel = Dlabel,
                              Xlabel = Xlabel, theme.bw = TRUE,CV = TRUE,
                              model.y = "nn",param.grid.y = param_grid_nn,
                              model.t = "hgb",param.grid.t = param_grid_hgb)

# rf (cv) + rf (cv) + rf (cv)
binary.cvrf.cvrf <- interflex(estimator="DML", data = d,
                              Y = Y, D = D, X = X, Z = Z,
                              Ylabel = Ylabel, Dlabel = Dlabel,
                              Xlabel = Xlabel, theme.bw = TRUE,CV = TRUE,
                              model.y = "rf",param.grid.y = param_grid_rf,
                              model.t = "rf",param.grid.t = param_grid_rf)

# hgb (cv) + hgb (cv) + hgb (cv)
binary.cvhgb.cvhgb <- interflex(estimator="DML",  data = d,
                              Y = Y, D = D, X = X, Z = Z,
                              Ylabel = Ylabel, Dlabel = Dlabel,
                              Xlabel = Xlabel, theme.bw = TRUE,CV = TRUE,
                              model.y = "hgb",param.grid.y = param_grid_hgb,
                              model.t = "hgb",param.grid.t = param_grid_hgb)
```

```{r, message=FALSE, warning=FALSE,cache=TRUE, fig.width=9,fig.height=9}

ggarrange(binary.cvnn.cvrf$figure,
          binary.cvnn.cvhgb$figure,
          binary.cvrf.cvrf$figure,
          binary.cvhgb.cvhgb$figure)
```

### Comparison of estimators

In this section, we compare the performance of DML estimator with previously introduced estimators.

```{r application1, message=FALSE, warning=FALSE,cache=TRUE}

out.dml.nn<-interflex(estimator='DML', data = d,
                      model.y = 'nn', model.t = 'nn',
                      Y=Y,D=D,X=X, Z = Z, treat.type = "discrete",
                      Xlabel=Xlabel, theme.bw = TRUE,
                      Ylabel=Ylabel, Dlabel=Dlabel,ylim=c(-0.4,1))

out.dml.rf<-interflex(estimator='DML', data = d, 
                      model.y = 'rf', model.t = 'rf', 
                      Y=Y,D=D,X=X, Z = Z, treat.type = "discrete",
                      Xlabel=Xlabel, theme.bw = TRUE,
                      Ylabel=Ylabel, Dlabel=Dlabel,ylim=c(-0.4,1))

out.dml.hgb<-interflex(estimator='DML', data = d, 
                       model.y = 'hgb', model.t = 'hgb', 
                       Y=Y,D=D,X=X, Z = Z, treat.type = "discrete",
                       Xlabel=Xlabel, theme.bw = TRUE,
                       Ylabel=Ylabel, Dlabel=Dlabel,ylim=c(-0.4,1))

out.aipw <- interflex(estimator = "lasso", Y=Y,D=D,X=X,data=d, 
                     Xlabel=Xlabel,Ylabel=Ylabel,ylim=c(-0.4,1), 
                     Dlabel=Dlabel, theme.bw = TRUE)

out.linear<-interflex(estimator = "binning",Y=Y,D=D,
                    X=X,Z=Z,data=d, Xlabel=Xlabel, 
                    Ylabel=Ylabel, Dlabel=Dlabel,
                    ylim=c(-0.4,1), theme.bw = TRUE)

out.kernel<-interflex(estimator = "kernel", data=d, Y=Y,D=D,X=X,Z=Z,
                      vartype = 'bootstrap', parallel = T,
                      Dlabel=Dlabel, Xlabel=Xlabel, theme.bw = TRUE,
                      Ylabel=Ylabel,ylim=c(-0.4,1))
```

```{r message=FALSE, warning=FALSE,cache=TRUE, fig.width=10,fig.height=10}
ggarrange(out.linear$figure, out.kernel$figure, out.aipw$figure,
          out.dml.nn$figure, out.dml.rf$figure, out.dml.hgb$figure, 
          common.legend = TRUE, legend = "none",
          labels = c("Binning", "Kernel", "AIPW-Lasso",
                     "DML: nn","DML: rf", "DML: hgb"))
```

The three DML estimator plots consistently show that the marginal effect of the threat on anger escalates with increasing partisan identity across all machine learning methods employed. However, these DML estimators appear to overfit when partisan identity is below 0.25, a region with sparse data representation. The widen confidence intervals in areas where partisan identity is lower (e.g., below 0.25) further indicate uncertainty in the model's estimations in these areas.

------------------------------------------------------------------------

## Continuous Treatment

In this section, we demonstrate the application of the DML estimator, focusing on scenarios with continuous treatment. We use the previous empirical example from @adiguzel2023out. The outcome is change in the vote share for the incumbent AKP party, treatment is change in congestion, measured by the number of patients per doctor at the nearest clinic, moderator is logarithm of median property value.

```{r}
load("data/Adiguzel2023a.RData")
d<-lograyic_models
d<-remove_labels(d)
d<-as.data.frame(d)
main<-"Adiguzel2023a"
D="Dwalk_3" 
Y="Dakp_3" 
X="lograyic09"

Z=c('Duniversity_3', 'Dodr_3', 'Dydr_3', 'Dpopulation_3', 'Dhospital_pri_3', 'Dhospital_pub_3')

Dlabel<-"Dwalk"
Ylabel<-"Dakp"
Xlabel<-"Property values (2009)"
```

### DML estimator

With continuous treatment, we need to specify the data type in `treat.type = continuous`. The estimator automatically plots both the pointwise and uniform CI.

```{r app2_nn_rf, message=FALSE, warning=FALSE,cache=TRUE, fig.width = 6, fig.height = 4.5}
con.dml<- interflex(estimator="dml", data = d,
                           Y = Y, D = D, X = X, Z = Z,
                           Ylabel = Ylabel, Dlabel = Dlabel,
                           Xlabel = Xlabel, theme.bw = TRUE,
                           treat.type = "continuous")
con.dml$figure
```

We can access the results of estimation through `est.dml`.

```{r}
head(con.dml$est.dml[[1]])
```

### Comparison of estimators

In this section, we compare the performance of DML estimator with previously introduced estimators.

```{r applicaton2, message=FALSE, warning=FALSE,cache=TRUE}

out.dml.nn<-interflex(estimator='DML', data = d,
                      Y = Y, D = D, X = X, Z = Z,
                      Ylabel = Ylabel, Dlabel = Dlabel,
                      Xlabel = Xlabel, theme.bw = TRUE,
                      model.y = 'nn', model.t = 'nn', 
                      ylim = c(-2,4), treat.type = "continuous")

out.dml.rf<-interflex(estimator='DML', data = d,
                      Y = Y, D = D, X = X, Z = Z,
                      Ylabel = Ylabel, Dlabel = Dlabel,
                      Xlabel = Xlabel, theme.bw = TRUE,
                      model.y = 'rf', model.t = 'rf', 
                      ylim = c(-2,4), treat.type = "continuous")

out.dml.hgb<-interflex(estimator='DML', data = d,
                      Y = Y, D = D, X = X, Z = Z,
                      Ylabel = Ylabel, Dlabel = Dlabel,
                      Xlabel = Xlabel, theme.bw = TRUE,
                      model.y = 'hgb', model.t = 'hgb', 
                      ylim = c(-2,4), treat.type = "continuous")

out.aipw <- interflex(estimator = "lasso", data = d,
                      Y = Y, D = D, X = X, Z = Z,
                      Ylabel = Ylabel, Dlabel = Dlabel,
                      Xlabel = Xlabel, theme.bw = TRUE,
                      ylim = c(-2,4), treat.type = "continuous")

out.est1<-interflex(estimator = "binning",data = d,
                      Y = Y, D = D, X = X, Z = Z,
                      Ylabel = Ylabel, Dlabel = Dlabel,
                      Xlabel = Xlabel, theme.bw = TRUE,
                      ylim = c(-2,4), treat.type = "continuous")

out.kernel<-interflex(estimator = "kernel", data = d,
                      Y = Y, D = D, X = X, Z = Z,
                      Ylabel = Ylabel, Dlabel = Dlabel,
                      Xlabel = Xlabel, theme.bw = TRUE,
                      ylim = c(-2,4), vartype = "bootstrap", 
                      treat.type = "continuous")

```

```{r app2_print, message=FALSE, warning=FALSE,cache=TRUE, fig.width=9,fig.height=9}
ggarrange(
          out.est1$figure, 
          out.kernel$figure,
          out.aipw$figure,
          out.dml.nn$figure, 
          out.dml.rf$figure, out.dml.hgb$figure, 
          common.legend = TRUE, legend = "none",
          labels = c("Binning", "Kernel","PO-Lasso", "DML: nn","DML: rf","DML: hgb"))
```

All the advance options, such as first stage model, parameter specification, and cross-validation for parameters are all available in the continuous treatment settings.

Here, we provide another example to explore how selecting different DML parameters affects performance. In this case, we specify the the number of trees in the forest (`n_estimators`), the minimum number of samples required to split an internal node (`min_samples_split`), and the the minimum number of samples required to be at a leaf node (`min_samples_leaf`), using the list `param_rf` and feed them to the DML estimator. We compare it with the DML estimator using \`rf\`\` as first stage model, but without adjusting parameters.

```{r app2_search, message=FALSE, warning=FALSE,cache=TRUE, fig.width = 9, fig.height = 5}
param_rf <- list(
  n_estimators = as.integer(200),
  min_samples_split = as.integer(5),
  min_samples_leaf = as.integer(4)
)


con.DML.rf.linear.para <- interflex(estimator="DML", data = d,
                              Y = Y, D = D, X = X, Z = Z,
                              Ylabel = Ylabel, Dlabel = Dlabel,
                              Xlabel = Xlabel, theme.bw = TRUE,
                              model.y = "rf", param.y = param_rf,
                              model.t = "rf", param.t = param_rf,
                              main = "Specified parameters",
                              treat.type = "continuous", ylim = c(-2,4))

out.dml.rf<-interflex(estimator='DML', data = d,
                      Y = Y, D = D, X = X, Z = Z,
                      Ylabel = Ylabel, Dlabel = Dlabel,
                      Xlabel = Xlabel, theme.bw = TRUE,
                      model.y = 'rf', model.t = 'rf', 
                      ylim = c(-2,4),
                      main = "Random forests")

ggarrange(out.dml.rf$figure,
          con.DML.rf.linear.para$figure)
```

<!-- ## Discrete outcome -->

<!-- Traditional parametric approaches for binary or count data typically rely on link functions—such as logit, probit, poisson, or negative binomial - to capture nonlinearity between predictors and the outcome. However, this introduces two key challenges. First, it becomes difficult to disentangle the nonlinearity imposed by the link function from the intrinsic nonlinearity of the CME. Second, when the treatment (D) is continuous, the derivative (\frac{\partial Y}{\partial D}) depends on both (X) and (D), requiring an additional averaging over (D) to recover a coherent measure of the CME: (\mathbb{E}\left[\left.\frac{\partial Y}{\partial D}\right\rvert X = x\right]). -->

<!-- Similar to AIPW, when treatment is binary the link function can be naturally incorporated into the DML estimator as well. When the treatment is continuous, DML estimator can serve as good approximations. -->

<!-- To estimate CME with discrete outcome using DML estimator, the syntax is exactly the same with continuous treatment settings: -->

<!-- ```{r} -->

<!-- d<-load("data/ET2023.RData") -->

<!-- d<-as.data.frame(pol) -->

<!-- d<-remove_labels(d) -->

<!-- main<-"Egerod & Tran 2023" -->

<!-- D="did"  -->

<!-- Y="seats_t_ind"  -->

<!-- X="abs_nom" -->

<!-- Z = c("sen", "hloga") -->

<!-- # Dlabel<-"Honest Leadership and Open Government Ac" -->

<!-- Dlabel<-"Treatment"  -->

<!-- # Ylabel<-"ATT of HLOGA on Board Service"  -->

<!-- Ylabel<-"Outcome"  -->

<!-- Xlabel<-"Ideological Extremism" -->

<!-- cl <- "state_po" -->

<!-- ``` -->

<!-- ```{r app_discrete, message=FALSE, warning=FALSE,cache=TRUE} -->

<!-- binary.dml.nn.nn <- interflex(estimator="DML", data = d, -->

<!--                               Y = Y, D = D, X = X, Z = Z, na.rm = TRUE, -->

<!--                               Ylabel = Ylabel, Dlabel = Dlabel, -->

<!--                               Xlabel = Xlabel, theme.bw = TRUE, -->

<!--                               model.y = "nn",model.t = "nn", -->

<!--                               main = "Neural Network") -->

<!-- ## random forest + random forest -->

<!-- binary.dml.rf.rf <- interflex(estimator="DML", data = d, -->

<!--                               Y = Y, D = D, X = X, Z = Z, -->

<!--                               Ylabel = Ylabel, Dlabel = Dlabel, -->

<!--                               Xlabel = Xlabel, theme.bw = TRUE, -->

<!--                               model.y = "rf",model.t = "rf", -->

<!--                               main = "Random Forest", na.rm = TRUE,) -->

<!-- ## hist gradient boosting + hist gradient boosting -->

<!-- binary.dml.hgb.hgb <- interflex(estimator="DML",data = d, -->

<!--                               Y = Y, D = D, X = X, Z = Z, -->

<!--                               Ylabel = Ylabel, Dlabel = Dlabel, -->

<!--                               Xlabel = Xlabel, theme.bw = TRUE, -->

<!--                               model.y = "hgb",model.t = "hgb", -->

<!--                               main = "Hist Gradient Boosting", -->

<!--                               na.rm = TRUE,) -->

<!-- ``` -->

<!-- ```{r message=FALSE, warning=FALSE,cache=TRUE, fig.width = 9, fig.height = 5} -->

<!-- ggarrange(binary.dml.nn.nn$figure,  -->

<!--           binary.dml.rf.rf$figure, -->

<!--           binary.dml.hgb.hgb$figure, ncol = 3) -->

<!-- ``` -->
