# Classic Approaches {#sec-classic}

In this chapter, we show how to use **interflex** package to analyze and estimate CME. We begin with visual diagnostics such as subgroup scatterplots, binning plot, and Generalized Additive Models (GAM) plot. We then estimate CME with classic approaches, including linear estimator and kernel estimator; We also show how to use **interflex** package with binning, linear and kernel estimators when users are interested in extension function, such as multiple treatment arms, fixed effect, prediction, and analyzing differences in treatment effects. We illustrates the implementation details using empirical example, with binary treatment settings and continuous treatment settings.

<!-- and 2. outcome-based approach, including IPW estimator, and AIPW estimator. . -->

## Binary treatment example

In this example and through out the whole manual, we will use a consistent notation. $Y$ is the outcome variable, the moderator is $X$, the covariates are $Z$.

The example we are using is from @huddy2015expressive, which explores the effect of the expressive model of partisanship. Drawing on four survey studies, the authors argue that partisan identity drives campaign participation and strong emotional responses to ongoing campaign events. The Conditional Marginal Effect (CME) studied in the paper suggests that strongly identified partisans are more likely to exhibit stronger emotional reactions – such as anger when threatened with electoral loss - compared to those with weaker partisan identities. The variables of interest include:

-   Outcome variable: level of anger (continuous scale from 0 to 1)

-   Treatment variable: electoral loss threat or electoral win reassurance (binary yes/no variable)

-   Moderator Variable: strength of partisan identity (continuous scale from 0 to 1)

```{r include=FALSE}
source("init.R")
```

```{r message = FALSE, warning = FALSE}
library(interflex)
data(interflex)
ls()
```

```{r app1_data, message=FALSE, warning=FALSE, cathe = TRUE}
d<-app_hma2015

Y = "totangry" ## outcome: anger
D = "threat" ## treatment: threat
X = "pidentity" ## moderator: partisan identity 
Z <-  c("issuestr2", "knowledge", "educ","male", "age10")

Ylabel = "Anger"
Dlabel = "Threat"
Xlabel = "Partisanship"

datatable(d)
```

------------------------------------------------------------------------

### Diagnostics

#### Raw plots

The first step of the diagnostics is to plot raw data. We supply the function **interflex** with the variable names of the outcome `Y`, the treatment `D`, and the moderator `X`, and covaraites `Z`. You can also supply labels for these variables, such as `Ylabel`, `Dlabel`, and `Xlabel`. If you supply a variable name to the `weights` option, the linear and LOESS fits will be adjusted based on the weights. Note that the correlations between covariates `Z` and `Y` are NOT partialled out.

<!-- We use `main` to add a title to the plot, `cex.main` to adjust its size, and `ylab` to change the y-axis label. -->

```{r binary_raw1, message=FALSE, warning=FALSE, cathe = TRUE, fig.width = 6, fig.height = 4.5}

interflex(estimator = "raw",data = d, Y = Y, 
                     D = D, X = X, Z = Z,
                     Ylabel = Ylabel, Dlabel = Dlabel,
                     Xlabel = Xlabel, theme.bw = TRUE,
          main = "Raw plot", ylab = "CME of Threat on Anger")
```

<!-- A black-white theme is applied when we set `theme.bw = TRUE`. `show.grid = FALSE` can be used to remove grid in the plot. Both options are allowed in **interflex** and **plot**. -->

<!-- ```{r} -->

<!-- interflex(estimator = "raw", data = d, Y = Y,  -->

<!--                      D = D, X = X, Z = Z, -->

<!--                      Ylabel = "Anger", Dlabel = "Threat", -->

<!--                      Xlabel = "Partisanship", -->

<!--           theme.bw = TRUE, show.grid = FALSE) -->

<!-- ``` -->

#### Binning plots

The second diagnostic tool is the binning plot. It will discretizes the moderator $X$ into several bins and create a dummy variable for each bin and picks an evaluation point within each bin, where users want to estimate the CME of $D$ on $Y$. The regression terms includes interactions between the bin dummies and the treatment indicator, the bin dummies and the moderator $X$ minus the evaluation points, as well as the triple interactions.

The histograms at the bottom of the plot depict the distributions of $X$ across treatment (pink) and control (gray) group. If the CME estimates from the binning estimator sit closely to the estimated CME from the linear estimator, then the Linear Interaction Effect (LIE) assumption likely holds

The `nbins` option sets the number of bins. The default number of bins is 3, and equal-sized bins are created based on the distribution of the moderator.

```{r binary_bin1, message=FALSE, warning=FALSE, cathe = TRUE, fig.width = 6, fig.height = 4.5}
bin.out.1 <- interflex(estimator = "binning", data = d, 
                 Y = Y,D = D, X = X, Z = Z,
                 Ylabel = Ylabel, Dlabel = Dlabel,
                 Xlabel = Xlabel,
                 theme.bw = TRUE,
                 main = "Marginal Effects")
bin.out.1$figure
```

For fully moderated model, we includes interactions between each covariate $Z$ and the bin dummies and the triple interactions. It can reduce biases when the impact of some covariates to the outcome does not take a linear form of moderator $X$, but sometimes it will also render the estimates unstable when the number of covariates is large.

```{r binary_bin2, message=FALSE, warning=FALSE, cathe = TRUE, fig.width = 6, fig.height = 4.5}
bin.out.2 <- interflex(estimator = "binning", data = d, 
                 Y = Y,D = D, X = X, Z = Z,
                 Ylabel = Ylabel, Dlabel = Dlabel,
                 Xlabel = Xlabel,
                 theme.bw = TRUE, full.moderate = TRUE,
                 main = "Marginal Effects")
bin.out.2$figure
```

Note that we can customize the cutoff values for the bins, for example, set `cutoffs = c(0.25,0.5,0.75,0.9)` to create five bins: $[minX, 0.25], (0.25, 0.5], (0.5,0.75], (0.75, 0.9]$ and $(0.9,maxX]$ (supplying N numbers will create N+1 bins). Note that the `cutoffs` option will override the **nbins** option if they are incompatible.

```{r binary_bin3, message=FALSE, warning=FALSE, cathe = TRUE, fig.width = 6, fig.height = 4.5}
cutoff<-c(0.25,0.5,0.75,0.9)

bin.out.3 <- interflex(estimator = "binning", data = d, 
                 Y = Y,D = D, X = X, Z = Z,
                 Ylabel = Ylabel, Dlabel = Dlabel,
                 Xlabel = Xlabel,
                 theme.bw = TRUE, nbins = 5,
                 main = "Marginal Effects")
bin.out.3$figure
```

Next, we use `Xunif = TRUE` to transform the moderator into a uniformly distributed random variable (based on the rank order in values of the orginal moderator) before estimating the CME.

```{r binary_bin4, message=FALSE, warning=FALSE, cathe = TRUE, fig.width = 6, fig.height = 4.5}
bin.out.4 <- interflex(estimator = "binning", data = d, 
                 Y = Y,D = D, X = X, Z = Z,
                 Ylabel = Ylabel, Dlabel = Dlabel,
                 Xlabel = Xlabel,
                 theme.bw = TRUE, Xunif = TRUE,
                 main = "Marginal Effects")
bin.out.4$figure
```

**interflex** will also automatically report a set of statistics when `estimator = "binning"`, including: (1) the binning estimates and their standard errors and 95% confidence intervals, (2) the percentage of observations within each bin, (3) the L-kurtosis of the moderator, and (4) a Wald test to formally test if we can reject the linear multiplicative interaction model by comparing it with a more flexible model of multiple bins.

The CME plot is stored in `object$figure` while the estimates and standard errors are stored in `object$est.bin` (binning). The tests results are stored in `object$tests` (binning).

```{r binary_bin5, message=FALSE, warning=FALSE, cathe = TRUE}
print(bin.out.1$tests$p.wald)
```

`plot` allows users to adjust graphic options without re-estimating the model. The first entry must be a **interflex** object. Note that we use `bin.labs = FALSE` to hide the label on the top of each bin and `Xdistr = "none` to remove the distribution of the moderator (not recommended), or set `Xdistr = "density` to present the distribution of the moderator with a density plot using option. We use `cex.axis` and `cex.lab` to adjust the font sizes of axis numbers and labels. We address all the plotting options in chapter @sec-plotting.

```{r binary_bin6, message=FALSE, warning=FALSE, cathe = TRUE, fig.width = 6, fig.height = 4.5}

plot(bin.out.1, Xdistr = "density", bin.labs = FALSE, theme.bw = TRUE)
```

------------------------------------------------------------------------

### Classic approaches

In this section, we discuss classic approaches to estimating and visualizing the CME. We begin with the linear interaction model, which remains widely used in social science research. To relax functional form assumptions, we then explore the kernel estimator proposed by HMX (2019).

#### Linear estimator

**Estimation.** If we want to deploy the linear estimator, we can set `estimator = "linear"`. This method estimates CME with multiplicative interaction model. When the option `full.moderate` is set to `TRUE`, interaction terms between the moderator $X$ and each covariate $Z$ will enter the model, which is the case in Blackwell and Olson (2020). This extension in some cases can reduce the bias of the original model.

**Uncertainty estimation.** Depending on `vartype`, the estimator computes standard errors and confidence intervals in three ways `simu`, `delta`, and `bootstrap`. The option `vartype = "simu"` draws from a multivariate normal distribution multiple times, `vartype = "delta"` uses a matrix derivatives to approximate the variance.

We report the uniform/simultaneous confidence interval along with pointwise confidence interval. Pointwise CI is estimated by default, and is depicted by gray ribbon; Uniform CI is estimated when setting `vartype = "bootstrap"` and `vartype = "delta"`, as depicted by the dashed lines.

Note that the variance-covariance matrix of coefficients used by the delta method or the simulation method by default is the robust standard error matrix estimated from the regression. Users can also choose `homoscedastic` or `cluster` standard error by specifying the option `vcov.type`. There are four options for the choice of the `vcov` estimator: `vcov.type = "homoscedastic"`, `"robust"`, `"cluster"`, and `"pcse"`. The default option is `"robust"`.

```{r binary_lin1, message=FALSE, warning=FALSE, cathe = TRUE, fig.width = 9, fig.height = 5}

linear.out.1 <- interflex(estimator = "linear", data = d, 
                 Y = Y,D = D, X = X, Z = Z,
                 Ylabel = Ylabel, Dlabel = Dlabel,
                 Xlabel = Xlabel, vartype = "bootstrap",
                 theme.bw = TRUE,
                 main = "Linear model")

linear.out.2 <- interflex(estimator = "linear", data = d, 
                 Y = Y,D = D, X = X, Z = Z,
                 Ylabel = Ylabel, Dlabel = Dlabel,
                 Xlabel = Xlabel,
                 vartype = "bootstrap",
                 theme.bw = TRUE, full.moderate = TRUE,
                 main = "Fully Moderated Linear model")

ggarrange(linear.out.1$figure, linear.out.2$figure)
```

**Results.** The CME plot is stored in `out$figure` while the estimates and standard errors are stored in `out$est.lin`

```{r binary_lin2, message=FALSE, warning=FALSE, cathe = TRUE}
head(linear.out.1$est.lin$`1`)
```

#### kernel estimator

**Estimation.** Similar to other estimator, to deploy the kernel estimator, we can set `estimator = "kernel"`. We also provide `full.moderate` option, allowing moderator $X$ and each covariate $Z$ to fully interact.

**Uncertainty estimation.** Kernel estimator computes standard errors in two ways: `delta` and `bootstrap`. Uniform CI is estimated when setting `vartype = "bootstrap"`.

When the standard errors are produced bootstrapping, we can adjust the number of bootstrap iterations using the `nboots` option, with the default being 200. The grid option can either be an integer, which specifies the number of bandwidths to be cross-validated, or a vector of candidate bandwidths – the default is `grid = 20`. We can also specify a clustered group structure by using the `cl` option, in which case a block bootstrap procedure will be performed.

The left panel uses the default settings for the kernel estimator; the right panel use fully interacted model, with `nboots = 400`, and we set the estimator to choose bandwidths from `c(0,1,2,3)`

```{r binary_kernel1, message=FALSE, warning=FALSE, cathe = TRUE, fig.width = 9, fig.height = 5}
kernel.out.1 <- interflex(estimator = "kernel", 
                          data = d, 
                 Y = Y,D = D, X = X, Z = Z,
                 Ylabel = Ylabel, Dlabel = Dlabel,
                 Xlabel = Xlabel,
                 vartype = "bootstrap", 
                 theme.bw = TRUE, bw = 0.916,
                 main = "Kernel model")


kernel.out.2 <- interflex(estimator = "kernel", 
                          data = d, 
                 Y = Y,D = D, X = X, Z = Z,
                 Ylabel = Ylabel, Dlabel = Dlabel,
                 Xlabel = Xlabel,
                 vartype = "bootstrap", grid = c(0,1,2,3),
                 theme.bw = TRUE, full.moderate = TRUE,
                 main = "Fully Moderated Kernel model")

ggarrange(kernel.out.1$figure, kernel.out.2$figure)
```

**Bandwidth selection.** We use adaptive bandwidth selection in **interflex**. This means that the bandwidth will be smaller in regions where there are more data and larger in regions where there are few data points, based on the moderator $X$. When adaptive bandwidth is being used, `bw` refers to the bandwidth applied in the region with the highest density of observations.

Even with an optimized algorithm, the bootstrap procedure for large datasets can be slow. We incorporate parallel computing, `parallel=TRUE`, to speed it up. You can choose the number of cores to be used for parallel computing; otherwise the program will detect the number of logical cores in your computer and use as many cores as possible (warning: this will significantly slow down your computer!).

If you specify a bandwidth manually, for example, by setting `bw = 1`, the cross-validation step will be skipped.

```{r binary_kernel2, message=FALSE, warning=FALSE, cathe = TRUE}
cat("1. Runtime with parallel computing for Bandwidth selection:\n")
runtime_parallel <- system.time({
  kernel.out.parallel <- interflex(estimator = "kernel", data = d, 
                                   Y = Y, D = D, X = X, Z = Z,
                                   Ylabel = Ylabel, Dlabel = Dlabel,
                                   Xlabel = Xlabel,
                                   vartype = "bootstrap", 
                                   theme.bw = TRUE, parallel = TRUE,
                                   main = "Linear model")
})
cat("User Time: ", runtime_parallel[1], " seconds\n")
cat("System Time: ", runtime_parallel[2], " seconds\n")
cat("Elapsed Time: ", runtime_parallel[3], " seconds\n\n")

cat("2. Runtime with user-specified Bandwidth selection:\n")
runtime_user_bandwidth <- system.time({
  kernel.out.user_bandwidth <- interflex(estimator = "kernel", data = d, 
                                         Y = Y, D = D, X = X, Z = Z,
                                         Ylabel = Ylabel, Dlabel = Dlabel,
                                         Xlabel = Xlabel,
                                         bw = 0.9,
                                         theme.bw = TRUE, 
                                         main = "Linear model")
})
cat("User Time: ", runtime_user_bandwidth[1], " seconds\n")
cat("System Time: ", runtime_user_bandwidth[2], " seconds\n")
cat("Elapsed Time: ", runtime_user_bandwidth[3], " seconds\n")
```

**Results.** The estimation results, including pointwise and uniform confidence intervals, are included in `est.kernel`.

```{r binary_kernel3, message=FALSE, warning=FALSE, cathe = TRUE}
head(kernel.out.1$est.kernel$`1`)
```

The results of cross-validation are saved in `CV.output`.

```{r binary_kernel4, message=FALSE, warning=FALSE, cathe = TRUE}
head(kernel.out.1$CV.output)
```

------------------------------------------------------------------------

## Continuous treatment example

@adiguzel2023out examines the impact of incumbent's public health provision on electoral outcomes. The authors use a reform in Turkey that significantly altered the geographic distribution of health clinics. The authors find that reduced congestion levels significantly increased the AKP's vote share, especially in poorer communities which exhibited a more substantial response to the improvements in health care access. The variables of interest are:

-   Outcome: Change in the vote share for the incumbent AKP party (continuous, $\in [-33.9, 39.5]$)
-   Treatment: Change in congestion (continuous, $\in [-28.8, 20.8]$); congestion is measured by the number of patients per doctor at the nearest clinic, with lower congestion indicating improved service quality
-   Moderator: Logarithm of median property value (continuous, $\in[-0.2, 7.7]$)

```{r con_data, message=FALSE, warning=FALSE, cathe = TRUE}
load("data/Adiguzel2023a.RData")
d<-lograyic_models
d<-remove_labels(d)
d<-as.data.frame(d)
main<-"Adiguzel2023a"
D="Dwalk_3" 
Y="Dakp_3" 
X="lograyic09"

Z=c('Duniversity_3', 'Dodr_3', 'Dydr_3', 'Dpopulation_3', 'Dhospital_pri_3', 'Dhospital_pub_3')

Dlabel<-"Dwalk"
Ylabel<-"Dakp"
Xlabel<-"Property values (2009)"

#datatable(d)
```

------------------------------------------------------------------------

### Diagnostics

#### Raw plots

Similar to the binary treatment settings, we can plot the subgroup scatter plot as first step of diagnostics:

```{r con_raw1, message=FALSE, warning=FALSE, cathe = TRUE, fig.width = 6, fig.height = 4.5}
interflex(estimator = "raw",data = d, Y = Y, 
                     D = D, X = X, Z = Z,
                     Ylabel = Ylabel, Dlabel = Dlabel,
                     Xlabel = Xlabel,
          main = "Raw plot", ylab = "CME", theme.bw = TRUE)
```

Instead of grouping by treatment status, the scatter plot carves $X$ into 3 quantile‐based bins by default. You can supply your own number of bins by setting the `cutoff =`. Supplying $N$ numbers will create $N+1$ bins.

```{r con_raw2, message=FALSE, warning=FALSE, cathe = TRUE, fig.width = 6, fig.height = 4.5}
cut_bins = c(3, 4, 5)
interflex(estimator = "raw",data = d, Y = Y, 
                     D = D, X = X, Z = Z,
                     Ylabel = Ylabel, Dlabel = Dlabel,
                     Xlabel = Xlabel, cutoff = cut_bins,
          main = "Raw plot", ylab = "CME", theme.bw = TRUE)
```

#### Binning plots

Similar to the binary treatment case, we can also use binning plot for diagnosis.

```{r con_bin1, message=FALSE, warning=FALSE, cathe = TRUE, fig.width = 6, fig.height = 4.5}
bin.out.1<-interflex(estimator = "binning",data = d, Y = Y, 
                     D = D, X = X, Z = Z,
                     Ylabel = Ylabel, Dlabel = Dlabel,
                     Xlabel = Xlabel, cutoff = cut_bins,
          main = "Binning plot", ylab = "CME", theme.bw = TRUE)

bin.out.1$figure
```

Because the treatment group is not binary, the histograms at the bottom of the plot simply depict the distribution of $X$, not differentiating by treatment status.

#### GAM plots

For the continuous treatment case, we can draw a Generalized Additive Model (GAM) plot.

```{r con_gam1, message=FALSE, warning=FALSE, cathe = TRUE,eval=FALSE,fig.width = 6, fig.height = 4.5}
interflex(estimator = "gam",data = d, Y = Y, 
                     D = D, X = X, Z = Z,
                     Ylabel = Ylabel, Dlabel = Dlabel,
                     Xlabel = Xlabel, show.uniform.CI = FALSE,
          main = "GAM plot", ylab = "CME", theme.bw = TRUE)
```

You can manipulate the angle of each of the four plots by setting `angle` with your own value of theta, which change the azimuth (horizontal rotation) angle in degrees. The default angle is `c(30, 100,-30,-120).`

```{r con_gam2, message=FALSE, warning=FALSE, cathe = TRUE, fig.width = 6, fig.height = 4.5}
interflex(estimator = "gam",data = d, Y = Y, 
                     D = D, X = X, Z = Z,
                     Ylabel = Ylabel, Dlabel = Dlabel,
                     Xlabel = Xlabel, angle = c(40, 50, 100, 120),
          main = "Raw plot", ylab = "CME", theme.bw = TRUE)
```

------------------------------------------------------------------------

### Classic approaches

In this section, we illustrate the code implementation using an empirical example with a continuous treatment. Note that the coding approach is essentially the same for both continuous and binary treatments

#### Linear estimator

```{r con_lin1, message=FALSE, warning=FALSE, cathe = TRUE, fig.width = 9, fig.height = 5}
linear.out.1 <- interflex(estimator = "linear", data = d, 
                 Y = Y,D = D, X = X, Z = Z,
                 Ylabel = Ylabel, Dlabel = Dlabel,
                 Xlabel = Xlabel,
                 theme.bw = TRUE,
                 main = "Linear model")

linear.out.2 <- interflex(estimator = "linear", data = d, 
                 Y = Y,D = D, X = X, Z = Z,
                 Ylabel = Ylabel, Dlabel = Dlabel,
                 Xlabel = Xlabel,
                 theme.bw = TRUE, full.moderate = TRUE,
                 main = "Fully Moderated Linear model")

ggarrange(linear.out.1$figure, linear.out.2$figure)
```

#### Kernel estimator

We can deploy kernel estimator with the same coding syntax.

```{r con_kernel1, message=FALSE, warning=FALSE, cathe = TRUE, fig.width = 9, fig.height = 5}
kernel.out.1 <- interflex(estimator = "kernel", 
                          data = d, 
                 Y = Y,D = D, X = X, Z = Z,
                 Ylabel = Ylabel, Dlabel = Dlabel,
                 Xlabel = Xlabel,
                 vartype = "bootstrap", 
                 theme.bw = TRUE, 
                 main = "Kernel model")

kernel.out.2 <- interflex(estimator = "kernel", 
                          data = d, 
                 Y = Y,D = D, X = X, Z = Z,
                 Ylabel = Ylabel, Dlabel = Dlabel,
                 Xlabel = Xlabel,
                 vartype = "bootstrap", grid = c(0,1,2,3),
                 theme.bw = TRUE, full.moderate = TRUE,
                 main = "Fully Moderated Kernel model")

ggarrange(kernel.out.1$figure, kernel.out.2$figure)
```

------------------------------------------------------------------------

## Extensions

In this section, we introduce several extending functions of the binning, linear, and kernel estimator. These extending functions do not depend on the data type of treatment, hence they are applicable to almost all settings. To adopt to the flexibility of these extending function, we use simulation data to illustrate the coding implementation.

```{r}
data(interflex)
ls()
```

s1 is a case of a dichotomous treatment indicator with linear conditional marginal effects (CME); s2 is a case of a continuous treatment indicator with linear CME; s3 is a case of a dichotomous treatment indicator with nonlinear CME; s4 is a case of a dichotomous treatment indicator, nonlinear CME, with additive two-way fixed effects; and s5 is a case of a discrete treatment indicator, nonlinear CME, with additive two-way fixed effects. s1-s5 are generated using the following code:

```{r}
set.seed(1234)
n<-200
d1<-sample(c(0,1),n,replace=TRUE) # dichotomous treatment
d2<-rnorm(n,3,1) # continuous treatment
x<-rnorm(n,3,1) # moderator
z<-rnorm(n,3,1) # covariate
e<-rnorm(n,0,1) # error term

## linear marginal effect
y1<-5 - 4 * x - 9 * d1 + 3 * x * d1 + 1 * z + 2 * e
y2<-5 - 4 * x - 9 * d2 + 3 * x * d2 + 1 * z + 2 * e
s1<-cbind.data.frame(Y = y1, D = d1, X = x, Z1 = z)
s2<-cbind.data.frame(Y = y2, D = d2, X = x, Z1 = z)

## quadratic marginal effect
x3 <- runif(n, -3,3) # uniformly distributed moderator
y3 <- d1*(x3^2-2.5) + (1-d1)*(-1*x3^2+2.5) + 1 * z + 2 * e
s3 <- cbind.data.frame(D=d1, X=x3, Y=y3, Z1 = z)

## adding two-way fixed effects
n  <- 500
d4 <-sample(c(0,1),n,replace=TRUE) # dichotomous treatment
x4 <- runif(n, -3,3) # uniformly distributed moderator
z4 <- rnorm(n, 3,1) # covariate
alpha <- 20 * rep(rnorm(n/10), each = 10)
xi <- rep(rnorm(10), n/10)
y4 <- d4*(x4^2-2.5) + (1-d4)*(-1*x4^2+2.5) + 1 * z4 + 
  alpha + xi + 2 * rnorm(n,0,1)
s4 <- cbind.data.frame(D=d4, X=x4, Y=y4, Z1 = z4, unit = rep(1:(n/10), each = 10), year = rep(1:10, (n/10)))

## Multiple treatment arms
n <- 600
# treatment 1
d1 <- sample(c('A','B','C'),n,replace=T)
# moderator
x <- runif(n,min=-3, max = 3)
# covriates
z1 <- rnorm(n,0,3)
z2 <- rnorm(n,0,2)
# error
e <- rnorm(n,0,1)
y1 <- rep(NA,n)
y1[which(d1=='A')] <- -x[which(d1=='A')]
y1[which(d1=='B')] <- (1+x)[which(d1=='B')]
y1[which(d1=='C')] <- (4-x*x-x)[which(d1=='C')]
y1 <- y1 + e + z1 + z2
s5 <- cbind.data.frame(D=d1, X=x, Y=y1, Z1 = z1,Z2 = z2)
```

**s6** is a case of a dichotomous treatment indicator with the DGP $$E(Y)=prob(Y=1)=logit^{-1}(-1+D+X+DX)=\frac{e^{-1+D+X+DX}}{1+e^{-1+D+X+DX}}$$, where D is binary too.

**s7** is a case of a continuous treatment indicator with the same DGP as **s6** except that D is continuous;

**s8** is a case of a dichotomous treatment indicator with nonlinear "linear predictor": $$E(Y)=prob(Y=1)=logit^{-1}(1+D+X+DX-DX^{2})$$

**s9** is a case of a discrete treatment indicator in which for the group 0, $E(Y)=logit^{-1}(-X)$, for the group 1 $E(Y)=logit^{-1}(1+X)$, and for the group 2, $E(Y)=logit^{-1}(-X^{2}-X+4)$.

```{r}
set.seed(110)
n <- 2000
x <- runif(n,min=-3, max = 3)
d1 <- sample(x = c(0,1),size = n,replace = T)
d2 <- runif(min = 0,max = 1,n = n)
d3 <- sample(x = c(0,1,2),size = n,replace = T)
Z1 <- runif(min = 0,max = 1,n = n)
Z2 <- rnorm(n,3,1)

## s6
link1 <- -1+d1+x+d1*x
prob1 <- exp(link1)/(1+exp(link1))
rand.u <- runif(min = 0,max = 1,n = n)
y1 <- as.numeric(rand.u<prob1)
s6 <- cbind.data.frame(X=x,D=d1,Z1=Z1,Z2=Z2,Y=y1)

## s7
link2 <- -1+d2+x+d2*x
prob2 <- exp(link2)/(1+exp(link2))
rand.u <- runif(min = 0,max = 1,n = n)
y2 <- as.numeric(rand.u<prob2)
s7 <- cbind.data.frame(X=x,D=d2,Z1=Z1,Z2=Z2,Y=y2)

## s8
link3 <- -1+d1+x+d1*x-d1*x*x
prob3 <- exp(link3)/(1+exp(link3))
rand.u <- runif(min = 0,max = 1,n = n)
y3 <- as.numeric(rand.u<prob3)
s8 <- cbind.data.frame(X=x,D=d1,Z1=Z1,Z2=Z2,Y=y3)

## s9
link4 <- rep(NA,n)
link4[which(d3==0)] <- -x[which(d3==0)]
link4[which(d3==1)] <- (1+x)[which(d3==1)]
link4[which(d3==2)] <- (4-x*x-x)[which(d3==2)]

prob4 <- exp(link4)/(1+exp(link4))
rand.u <- runif(min = 0,max = 1,n = n)
y4 <- as.numeric(rand.u<prob4)
s9 <- cbind.data.frame(X=x,D=d3,Z1=Z1,Z2=Z2,Y=y4)
```

------------------------------------------------------------------------

### Fixed effects

First, let's look at the linear fixed effects models. Remember in **s4**, a large chunk of the variation in the outcome variable is driven by group fixed effects. Below is a scatterplot of the raw data (group index vs. outcome). Red and green dots represent treatment and control units, respectively. We can see that outcomes are highly correlated within a group.

```{r fe1, cache=TRUE, message=FALSE, fig.width = 6, fig.height = 4.5}
ggplot() + 
  geom_point(data = s4,aes(x = unit, y = Y, colour = as.factor(D))) + 
  guides(colour=FALSE) + theme_bw()
```

When fixed effects are present, it is possible that we cannot observe a clear pattern of CME in the raw plot as before, while binning estimates have wide confidence intervals:

```{r fe2a, results = "hide", cache=TRUE, message=FALSE, fig.width = 6, fig.height = 4.5}
interflex(estimator = "raw", Y = "Y", D = "D", X = "X", data = s4, 
          weights = NULL,ncols=2, theme.bw = TRUE)
```

```{r fe3, results = "hide", cache=TRUE, message=FALSE, fig.width = 6, fig.height = 4.5}
s4.binning <- interflex(Y = "Y", D = "D", X = "X", Z = "Z1", data = s4, 
                        estimator = "binning", FE = NULL, cl = "unit",
                        theme.bw = TRUE)
s4.binning$figure
```

The binning estimates are much more informative when fixed effects are included, by using the `FE` option. Note that the number of group indicators can exceed 2. Our algorithm is optimized for a large number of fixed effects or many group indicators. The `cl` option controls the level at which standard errors are clustered.

```{r fe4, results = "hide", cache=TRUE, message=FALSE, fig.width = 6, fig.height = 4.5}
s4$wgt <- 1
s4.binning <- interflex(Y = "Y", D = "D", X = "X", Z = "Z1", data = s4, 
                        estimator = "binning", FE = c("unit", "year"), 
                        cl = "unit", weights = "wgt", theme.bw = TRUE)

s4.binning$figure
```

When fixed effects are not taken into account, the kernel estimates are also less precisely estimated. Because the model is incorrectly specified, cross-validated bandwidths also tend to be bigger than optimal.

```{r fe5, results = "hide", cache=TRUE, message=FALSE, fig.width = 6, fig.height = 4.5}
s4.kernel <- interflex(Y = "Y", D = "D", X = "X", Z = "Z1", data = s4, 
                       estimator = "kernel", FE = NULL, vartype = 'bootstrap',
                       nboots = 200, parallel = TRUE, 
                       cores = 10,
                       cl = "unit", weights = "wgt", theme.bw = TRUE)
s4.kernel$figure
```

Controlling for fixed effects by using the `FE` option solves this problem. The estimates are now much closer to the population truth. Note that all uncertainty estimates are produced via bootstrapping. When the `cl` option is specified, a block bootstrap procedure will be performed.

```{r fe6, results = "hide", cache=TRUE, message=FALSE, fig.width = 6, fig.height = 4.5}
s4.kernel <- interflex(Y = "Y", D = "D", X = "X", Z = "Z1", data = s4, 
                       estimator = "kernel", FE = c("unit","year"), cl = "unit"
                       ,vartype = 'bootstrap', theme.bw = TRUE,
                       nboots = 200, parallel = TRUE, cores = 10,weights = "wgt")
s4.kernel$figure
```

With large datasets, cross-validation or bootstrapping can take a while. One way to check the result quickly is to shut down the bootstrap procedure (using `CI = FALSE`). **interflex** will then present the point estimates only. Another way is to supply a reasonable bandwidth manually by using the `bw` option such that cross-validation will be skipped.

```{r fe7, results = "hide", cache=TRUE, message=FALSE, fig.width = 6, fig.height = 4.5}
s4.kernel <- interflex(Y = "Y", D = "D", X = "X", Z = "Z1", data = s4, 
                       estimator = "kernel", bw = 0.62, FE = c("unit","year"), cl = "unit", 
                       vartype = 'bootstrap',
                       nboots = 200, parallel = TRUE, cores = 10,
                       CI = FALSE, 
                       ylim = c(-9, 15), theme.bw = TRUE)
s4.kernel$figure
```

Note that discrete outcome with fixed effects is not supported in the package.

------------------------------------------------------------------------

### Multiple treatment arms

Next, we will show how **interflex** can be applied when there are more than two condition. First, we plot the raw data:

```{r mta_raw, results = "hide", cache=TRUE, message=FALSE, fig.height=5, fig.width=10}
interflex(estimator = "raw",Y = "Y", D = "D", X = "X", data = s5, ncols = 3,
          theme.bw = TRUE)
```

By default, **interflex** will produce $(n-1)$ CME plots, taking one category as the baseline (if not specified, the base group is selected based on the numeric/character order of the treatment values). Here we also specify `vartype = "bootstrap"` in order to estimate the confidence intervals.

```{r mta1a, results = "hide", cache=TRUE, message=FALSE, fig.height=5, fig.width=8}
s5.binning <- interflex(Y = "Y", D = "D", X = "X", Z = c("Z1","Z2"), data = s5, 
                        estimator = "binning", vartype = "bootstrap",
                        theme.bw = TRUE)

s5.binning$figure
```

<!-- By setting the option `order` and `subtitles` (in either **interflex** or **plot**), users can also specify the order of plots they prefer and the subtitles they want:   -->

<!-- ```{r mta1c, message = FALSE, fig.height=5, fig.width=8, cache = TRUE,results = "hide"} -->

<!-- plot(s5.binning, order=c("C", "B"), subtitles = c("Group C","Group B"), theme.bw = TRUE) -->

<!-- ``` -->

<!-- By setting `pool = TRUE` (in either **interflex** or **plot**), we can combine these plots. Users can also specify the colors they like, :  -->

<!-- ```{r mta1d, message = FALSE, fig.height=5, fig.width=6, cache = TRUE} -->

<!-- plot(s5.binning, order=c("C", "B"), subtitles = c("Control Group","Group C","Group B"),  -->

<!--      pool = TRUE, color = c("Salmon","#7FC97F", "#BEAED4"), theme.bw = TRUE) -->

<!-- ``` -->

We specify the base group using the option `base`:

```{r mta2, cache=TRUE, results = "hide", message=FALSE, fig.height=5, fig.width=8}
s5.binning2 <- interflex(Y = "Y", D = "D", X = "X",Z = c("Z1", "Z2"), data = s5, 
                         estimator = "binning", base = "B", vartype = "bootstrap")
plot(s5.binning2, theme.bw = TRUE)
```

For dataset **s5**, because of the nonlinear interaction effects, the kernel estimator will produce CME estimates that are much closer to the effects implied by the true DGP.

```{r mta3, cache=TRUE, results = "hide", message=FALSE, fig.width = 6, fig.height = 4.5}
s5.kernel <- interflex(Y = "Y", D = "D", X = "X",Z = c("Z1", "Z2"), data = s5, 
                       estimator = "kernel", vartype = 'bootstrap',
                       nboots = 200, parallel = TRUE, cores = 10)
s5.kernel$figure
```

The function can be applied to discrete outcome settings. For **s9**, the LIE assumption are violated only in group 2. The linear estimator can give a consistent estimation of treatment effects for the group 1, but not group 2.

```{r dmulti1, cache=TRUE, results = "hide", message=FALSE, fig.height=5, fig.width=8}
s9.linear <- interflex(estimator='linear', data=s9, method='logit', 
                       Y = 'Y', D = 'D', X = 'X', Z = c('Z1','Z2'))
plot(s9.linear)
```

```{r dmulti2, cache=TRUE, results = "hide", message=FALSE, fig.height=5, fig.width=8}
TE1 <- exp(1+x)/(1+exp(1+x))-exp(-x)/(1+exp(-x))
p1 <- plot(s9.linear,show.all = T)$`1`+geom_line(aes(x=x,y=TE1),color='red')
TE2 <- exp(4-x*x-x)/(1+exp(4-x*x-x))-exp(-x)/(1+exp(-x))
p2 <- plot(s9.linear,show.all = T)$`2`+geom_line(aes(x=x,y=TE2),color='red')
p1 + p2
```

Then, we use the binning estimator to conduct a diagnostic analysis.

```{r dmulti3, cache=TRUE, results = "hide", message=FALSE, fig.height=5, fig.width=8}
s9.binning <- interflex(estimator = 'binning', data = s9, method='logit',
                        Y = 'Y', D = 'D', X = 'X', Z = c('Z1','Z2'))
plot(s9.binning)
```

Results of the binning estimator also suggests the violation of LIE for group 2, which can be further verified by checking the test results.

```{r dmulti4, cache=TRUE}
s9.binning$tests$p.wald
s9.binning$tests$p.lr
```

Both of these tests suggest that LIE assumption fails in the model, but it **doesn't** tell us in which group the assumption fails. There is an extra output "sub.test" when there are more than 3 groups in the data, which includes by-group results of the Wald/Likelihood ratio test, which test the linear multiplicative interaction assumption for each group. In this case, the test results suggest that there are not enough evidences to reject the null hypothesis for group 1, but for group 2, the null hypothesis can be safely rejected.

```{r dmulti5, cache=TRUE}
head(s9.binning$tests$sub.test[[1]])
```

Lastly, we can use the kernel estimator to get more consistent results.

```{r dmulti6, cache=TRUE, results = "hide", message=FALSE, fig.width = 6, fig.height = 4.5}
s9.kernel <- interflex(estimator = 'kernel', data = s9, 
                       Y ='Y', D = 'D', X = 'X', Z = c('Z1','Z2'),
                       method='logit',vartype = "bootstrap",
                      nboots = 200, parallel = TRUE, cores = 10)
plot(s9.kernel)
```

```{r dmulti7, cache=TRUE, results = "hide", message=FALSE, fig.height=5, fig.width=8}
TE1 <- exp(1+x)/(1+exp(1+x))-exp(-x)/(1+exp(-x))
p1 <- plot(s9.kernel,show.all = T)$`1`+geom_line(aes(x=x,y=TE1),color='red')
TE2 <- exp(4-x*x-x)/(1+exp(4-x*x-x))-exp(-x)/(1+exp(-x))
p2 <- plot(s9.kernel,show.all = T)$`2`+geom_line(aes(x=x,y=TE2),color='red')
p1 + p2
```

The estimated ATE and $E(Y|D,X,Z)$ are also grouped by treatment.

```{r dmulti8, cache=TRUE}
s9.kernel$Avg.estimate
```

```{r dmulti9, cache=TRUE, results = "hide", message=FALSE, fig.width = 8, fig.height = 4.5}
predict(s9.kernel)
```

------------------------------------------------------------------------

### Predicted outcomes

As we have shown earlier with discrete outcome settings, **interflex** allows users to estimate and visualize predicted outcomes given fixed values of the treatment and moderator based on flexible interaction models. All predictions are made by setting covariates equal to their sample means. The same functino applies to continuous outcome as well.

```{r ip3, fig.height=5, fig.width=8, cache=TRUE, results = "hide", message=FALSE}
s5.kernel <-interflex(Y = "Y", D = "D", X = "X", Z = c("Z1", "Z2"),
                      data = s5, estimator = "kernel",vartype = 'bootstrap',
                       nboots = 200, parallel = TRUE, cores = 10)
predict(s5.kernel, theme.bw = TRUE)
```

<!-- We can also set `pool = TRUE` to combine them. -->

<!-- ```{r ip5, fig.height=5, fig.width=8, cache=TRUE, results = "hide", message=FALSE} -->

<!-- predict(s5.kernel,order = c('A','B','C'),subtitle = c("Group A", "Group B", "Group C"),  -->

<!--         pool = TRUE, legend.title = "Three Different Groups") -->

<!-- ``` -->

We can also use the linear model or the binning model to estimate and visualize predicted outcomes. For the latter, it is expected to see some "zigzags" at the bin boundaries in a plot:

```{r ip6, fig.height=5, fig.width=8, cache=TRUE, results = "hide", message=FALSE}
s5.binning <-interflex(Y = "Y", D = "D", X = "X", Z = c("Z1","Z2"), data = s5, 
                       estimator = "binning", vartype = "bootstrap", nbins = 4)
predict(s5.binning, theme.bw = TRUE)
```

```{r ip7, fig.height=5, fig.width=8, cache=TRUE, results = "hide", message=FALSE}
s5.linear <-interflex(Y = "Y", D = "D", X = "X", Z = c("Z1","Z2"), 
                      data = s5, 
                      estimator = "linear", vcov.type = "robust")
predict(s5.linear, theme.bw = TRUE)
```

Obviously, the kernel estimator characterizes the data more better than a linear or binning estimator.

------------------------------------------------------------------------

### Differences in treatment effects

**interflex** allows users to compare treatment effects at three specific values of the moderator using linear or kernel model. In order to do so, users need to pass three values into `diff.values` when applying *linear* or *kernel* estimator. If not specified, this programm will by default choose the 25,50,75 percentile of the moderator as the `diff.values`.

```{r dbt1, fig.height=5, fig.width=10, message=FALSE,results = "hide", cache = TRUE}
s5.kernel <-interflex(Y = "Y", D = "D", X = "X", Z = c("Z1", "Z2"), data = s5, 
                      estimator = "kernel", diff.values = c(-2,0,2),
                      vartype = "bootstrap",
                      nboots = 200, parallel = TRUE, cores = 10,)
plot(s5.kernel,diff.values = c(-2,0,2), theme.bw = TRUE)
```

The difference of marginal effects/treatment effects at different values of the moderator(by default are the 25%, 50%, 75% quantiles of the moderator) are saved in `diff.estimate`.

```{r dbt2, message=FALSE}
s5.kernel$diff.estimate
```

```{r dbt3, fig.height=5, fig.width=10, message=FALSE,results = "hide", cache = TRUE}
s5.linear <-interflex(Y = "Y", D = "D", X = "X", Z = c("Z1","Z2"), data = s5, 
                      estimator = "linear", diff.values = c(-2,0,2), 
                      vartype = "bootstrap")
plot(s5.linear,diff.values = c(-2,0,2), theme.bw = TRUE)
```

```{r dbt4, message=FALSE}
s5.linear$diff.estimate
```

**interflex** allows users to compare treatment effects at two or three specific values of the moderator using CME and vcov matrix derived from linear/kernel estimation. Based on GAM model(relies on **mgcv** package), users can approximate treatment effects and their variance using smooth functions without re-estimating the model, hence saving time. In order to do so, users need to pass the output of **interflex** into **t.test** and then specify the values of interest of the moderator. As it is an approximation, the results here are expected to have a little deviance from the output we got by directly specifying `diff.values` when applying **inter.linear** or **inter.kernel**. We also limit the elements passed in diff.values between the minimum and the maximum of the moderator to avoid potential extrapolation bias.

```{r dbt5, message=FALSE, cache = TRUE}
inter.test(s5.kernel,diff.values=c(-2,0,2))
```

```{r dbt6, message=FALSE, cache = TRUE}
inter.test(s5.linear,diff.values=c(-2,0,2))
```

We can also set `percentile = TRUE` to use the percentile scale rather than the real scale of the moderator.

```{r dbt7, message=FALSE, cache = TRUE}
inter.test(s5.kernel,diff.values=c(0.25,0.5,0.75),percentile=TRUE)
```
