---
title: 'Interflex with Continuous Outcomes'
output:
  rmarkdown::html_vignette:
    toc: false
    toc_depth: 4
    number_sections: false
vignette: >
  %\VignetteIndexEntry{Interflex with Continuous Outcomes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}  
---
<!-- 
  Code to Justify Text
    <style>
    body {
    text-align: justify}
    </style>
-->   
```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache =TRUE)
knitr::opts_chunk$set(warning=FALSE)
``` 

Let's load the package as well as the simulated toy datasets. Note that **s6**-**s9** are for the tutorial for **interflex** with [discrete](discrete.html) outcomes.
```{r, message = FALSE}
library(interflex)
data(interflex)
ls()
```

**s1** is a case of a dichotomous treatment indicator with linear marginal effects; **s2** is a case of a continuous treatment indicator with linear marginal effects; **s3** is a case of a dichotomous treatment indicator with *nonlinear* marginal effects; **s4** is a case of a dichotomous treatment indicator, nonlinear marginal effects, with additive two-way fixed effects; and  **s5** is a case of a *discrete* treatment indicator, *nonlinear* marginal effects, with additive two-way fixed effects. **s1**-**s5** are generated using the following code:
```{r, eval = FALSE}
set.seed(1234)
n<-200
d1<-sample(c(0,1),n,replace=TRUE) # dichotomous treatment
d2<-rnorm(n,3,1) # continuous treatment
x<-rnorm(n,3,1) # moderator
z<-rnorm(n,3,1) # covariate
e<-rnorm(n,0,1) # error term

## linear marginal effect
y1<-5 - 4 * x - 9 * d1 + 3 * x * d1 + 1 * z + 2 * e
y2<-5 - 4 * x - 9 * d2 + 3 * x * d2 + 1 * z + 2 * e
s1<-cbind.data.frame(Y = y1, D = d1, X = x, Z1 = z)
s2<-cbind.data.frame(Y = y2, D = d2, X = x, Z1 = z)

## quadratic marginal effect
x3 <- runif(n, -3,3) # uniformly distributed moderator
y3 <- d1*(x3^2-2.5) + (1-d1)*(-1*x3^2+2.5) + 1 * z + 2 * e
s3 <- cbind.data.frame(D=d1, X=x3, Y=y3, Z1 = z)

## adding two-way fixed effects
n  <- 500
d4 <-sample(c(0,1),n,replace=TRUE) # dichotomous treatment
x4 <- runif(n, -3,3) # uniformly distributed moderator
z4 <- rnorm(n, 3,1) # covariate
alpha <- 20 * rep(rnorm(n/10), each = 10)
xi <- rep(rnorm(10), n/10)
y4 <- d4*(x4^2-2.5) + (1-d4)*(-1*x4^2+2.5) + 1 * z4 + 
  alpha + xi + 2 * rnorm(n,0,1)
s4 <- cbind.data.frame(D=d4, X=x4, Y=y4, Z1 = z4, unit = rep(1:(n/10), each = 10), year = rep(1:10, (n/10)))

## Multiple treatment arms
n <- 600
# treatment 1
d1 <- sample(c('A','B','C'),n,replace=T)
# moderator
x <- runif(n,min=-3, max = 3)
# covriates
z1 <- rnorm(n,0,3)
z2 <- rnorm(n,0,2)
# error
e <- rnorm(n,0,1)
y1 <- rep(NA,n)
y1[which(d1=='A')] <- -x[which(d1=='A')]
y1[which(d1=='B')] <- (1+x)[which(d1=='B')]
y1[which(d1=='C')] <- (4-x*x-x)[which(d1=='C')]
y1 <- y1 + e + z1 + z2
s5 <- cbind.data.frame(D=d1, X=x, Y=y1, Z1 = z1,Z2 = z2)
```

The implied population marginal effects for the DGPs of **s1** and **s2** are $$ME(X) = 3X - 9;$$ the implied population marginal effects for the DGPs of **s3** and **s4** are $$ME(X) = 2X^{2} - 5.$$ For **s5**, if we set treatment "A" as our base category, the implied population marginal effects for group "B" and group "C" are, respectively, $$ME(X) = 2X + 1\quad\text{and}\quad ME(X) = -X^{2} + 4.$$ 

The **interflex** package ships the following functions: **interflex**, **inter.raw**, **inter.gam**,  **plot**.  The functionalities of **inter.binning** and **inter.kernel** covered by **interflex**, but they are still supported for backward compatibility.

---

## Raw plots

The first step of the diagnostics is to plot raw data. We supply the function **interflex** with the variable names of the outcome  `Y`, the treatment `D`, and the moderator `X`. You can also supply labels for these variables. If you supply a variable name to the `weights` option, the linear and LOESS fits will be adjusted based on the weights. Note that the correlations between covariates `Z` and `Y` are NOT partialed out. We use `main` to add a title to the plot and `cex.main` to adjust its size.
```{r raw1, results = "hide", fig.height=4, fig.width=8}
interflex(estimator = "raw",Y = "Y", D = "D", X = "X", data = s1, 
          weights = NULL, Ylabel = "Outcome", 
          Dlabel = "Treatment", Xlabel="Moderator", 
          main = "Raw Plot", cex.main = 1.2, ncols=2)
```

A black-white theme is applied when we set `theme.bw = TRUE`. `show.grid = FALSE` can be used to remove grid in the plot. Both options are allowed in **interflex** and **plot**.
```{r raw2, results = "hide", fig.height=4, fig.width=10}
interflex(estimator = "raw", Y = "Y", D = "D", X = "X", data = s2, 
          Ylabel = "Outcome", Dlabel = "Treatment", Xlabel="Moderator", 
          theme.bw = TRUE, show.grid = FALSE, ncols=3)
```

```{r raw3, results = "hide", fig.height=4, fig.width=8}
interflex(estimator = "raw", Y = "Y", D = "D", X = "X", data = s3, 
          Ylabel = "Outcome", Dlabel = "Treatment", Xlabel="Moderator",
          ncols=3)
```

For the continuous treatment case (e.g. **s2**), we can also draw a Generalized Additive Model (GAM) plot. You can supply a set of covariates to be controlled for by supplying `Z`, which takes a vector of covariate names (strings). 
```{r gam, fig.height=5, fig.width=5, message=FALSE}
interflex(estimator = "gam",Y="Y", D="D", X="X", Z=c("Z1"), data=s2)
```
---

## The binning estimator

The second diagnostic tool is the binning plot. The `nbins` option sets the number of bins. The default number of bins is 3, and equal-sized bins are created based on the distribution of the moderator. There are four options for the choice of the vcov estimator: `vcov.type = "homoscedastic"`, `"robust"`, `"cluster"`,and `"pcse"`. The default option is `"robust"`.

Note that **interflex** will also automatically report a set of statistics when `estimator = "binning"`, including: (1) the binning estimates and their standard errors and 95% confidence intervals, (2) the percentage of observations within each bin, (3) the L-kurtosis of the moderator, and (4) a Wald test to formally test if we can reject the linear multiplicative interaction model by comparing it with a more flexible model of multiple bins. . 

```{r binning1aa, message = FALSE, fig.height=5, fig.width=5, cathe = TRUE}
out <- interflex(Y = "Y", D = "D", X = "X", Z = "Z1", data = s1, 
                 estimator = "binning", vcov.type = "robust", 
                 main = "Marginal Effects", ylim = c(-15, 15))
plot(out)
print(out$tests$p.wald)
```

We see that the Wald test cannot reject the NULL hypothesis that the linear interaction model and the three-bin model are statistically equivalent. If we only want to conduct the linear estimator, we can set `estimator = "linear"`:
```{r binning1ab, message = FALSE, fig.height=5, fig.width=5, cache = TRUE}
out2 <- interflex(Y = "Y", D = "D", X = "X", Z = "Z1", data = s1, 
                  estimator = "linear", vcov.type = "robust", 
                  main = "Marginal Effects", ylim = c(-15, 15))
plot(out2)
```

**plot** allows users to adjust graphic options without re-estimating the model. The first entry must be a `interflex` object. Note that we use `bin.labs = FALSE` to hide the label on the top of each bin and `Xdistr = "none"` to remove the distribution of the moderator (not recommended). We use `cex.axis` and `cex.lab` to adjust the font sizes of axis numbers and labels.
```{r binning1a, fig.height=5, fig.width=5, message=FALSE, warning=FALSE, paged.print=FALSE,results="hide"}
plot(out, xlab = "Moderate is X", Xdistr = "none", bin.labs = FALSE, cex.axis = 0.8, cex.lab = 0.8)
```

Next, we use `Xunif = TRUE` to transform the moderator into a uniformly distributed random variable (based on the rank order in values of the orginal moderator) before estimating the marginal effects. `nbins = 4` sets the number of bins to 4.
```{r binning1b, message = FALSE, fig.height=5, fig.width=5}
out <- interflex(Y = "Y", D = "D", X = "X", Z = "Z1", data = s1, 
                 estimator = "binning", nbins = 4, 
                 theme.bw = TRUE, Xunif = TRUE)
out$figure
```

The binning estimates for the continuous case are shown below. We now present the distribution of the moderator with a density plot using option `Xdist = "density"` -- the default option is `"hist"` or `"histogram"`. We turn off the bin labels using `bin.labs = FALSE`.
```{r binning2,  fig.height=5, fig.width=5, message = FALSE, cache = TRUE}
out <- interflex(Y = "Y", D = "D", X = "X", Z = "Z1", data = s2, 
                 estimator = "binning", Xdistr = "density", 
                 bin.labs = FALSE)
out$figure
```

Note that you can customize the cutoff values for the bins, for example, set `cutoffs = c(1, 2, 4, 5)` to create five bins: [minX, 1], (1, 2], (2,4], (4, 5] and (5,maxX] (supplying N numbers will create N+1 bins). Note that the `cutoffs` option will override the **nbins** option if they are incompatible. 
```{r binning3,  fig.height=5, fig.width=5, message = FALSE, cache = TRUE}
out <- interflex(Y = "Y", D = "D", X = "X", Z = "Z1", data = s2, 
                 estimator = "binning", cutoffs = c(1,2,4,5))
out$figure
```

The binning estimates for the dichotomous, nonlinear case (i.e. **s3**) are shown below. A linear interaction model clearly gives misleading marginal effects estimates. The marginal effects plot is stored in `out$figure` while the estimates and standard errors are stored in `out$est.linear` (linear) and `out$est.bin` (binning). The tests results are stored in `out$tests` (binning). 
```{r binning4, fig.height=5, fig.width=5, message = FALSE, cache = TRUE}
out <- interflex(Y = "Y", D = "D", X = "X", Z = "Z1", data = s3, 
                 estimator = "binning")
print(out$tests$p.wald)
```
This time the NULL hypothesis that the linear interaction model and the three-bin model are statistically equivalent is safely rejected (p.wald = 0.00).

---

## The kernel estimator

With the kernel method, a bandwidth is first selected via 10-fold least-squares cross-validation. 

The standard errors are produced by a non-parametric bootstrap (you can adjust the number of bootstrap iterations using the `nboots` option). The `grid` option can either be an integer, which specifies the number of bandwidths to be cross-validated, or a vector of candidate bandwidths -- the default is `grid = 20`. You can also specify a clustered group structure by using the `cl` option, in which case a block bootstrap procedure will be performed.

Starting from v.1.1.0, we introduce adaptive bandwidth selection to **interflex**. This means that the bandwidth will be smaller in regions where there are more data and larger in regions where there are few data points (based on the moderator). When adaptive bandwidth is being used, `bw` refers to the bandwidth applied in the region with the highest density of observations.

Even with an optimized algorithm, the bootstrap procedure for large datasets can be slow. We incorporate parallel computing (`parallel=TRUE`) to speed it up. You can choose the number of `cores` to be used for parallel computing; otherwise the program will detect the number of logical cores in your computer and use as many cores as possible (warning: this will significantly slow down your computer!). 

```{r kernel1, fig.height=5, fig.width=5, message=FALSE, cache=TRUE}
out <- interflex(Y = "Y", D = "D", X = "X", Z = "Z1", data = s1, 
                 estimator = "kernel", nboots = 200, 
                 parallel = TRUE, cores = 4)
out$figure
```

If you specify a bandwidth manually (for example, by setting `bw = 1`), the cross-validation step will be skipped. The `main` option controls the title of the plot while the `xlim` and `ylim` options control the ranges of the x-axis and y-axis, respectively.
```{r kernel2, cache=TRUE, message=FALSE, results = "hide", fig.height=5, fig.width=5}
out <- interflex(Y = "Y", D = "D", X = "X", Z = "Z1", data = s1, 
                 estimator = "kernel", nboots = 200, bw = 1, 
                 main = "The Linear Case", 
                 xlim = c(-0.5,6.5), ylim = c(-15, 15))
out$figure
```

For dataset **s3**, the kernel method produces non-linear marginal effects estimates that are much closer to the effects implied by the true DGP.
```{r kernel3, cache=TRUE, results = "hide", message=FALSE, fig.height=5, fig.width=5}
out <- interflex(Y = "Y", D = "D", X = "X", Z = "Z1", data = s3, 
                 estimator = "kernel", theme.bw = TRUE)
out$figure
```

Again, if we only want to change the look of a marginal effect plot, we can use **plot** to save time. `CI = FALSE` removes the confidence interval ribbon. 
```{r kernel3a, cache=TRUE, fig.height=5, fig.width=5,results = "hide"}
plot(out, main = "Nonlinear Marginal Effects", 
     ylab = "Coefficients", Xdistr = "density", 
     xlim = c(-3,3), ylim = c(-10,12), CI = FALSE, 
     cex.main = 0.8, cex.lab = 0.7, cex.axis = 0.7)
```

The semi-parametric marginal effect estimates are stored in `out$est.kernel`.

Note that we can use the `file` option to save a plot to a file in **interflex** (e.g. by setting `file = "myplot.pdf"` or `file = "myplot.png"`).

---

## Fixed effects

We move on to linear fixed effects models. Remember in **s4**, a large chunk of the variation in the outcome variable is driven by group fixed effects. Below is a scatterplot of the raw data (group index vs. outcome). Red and green dots represent treatment and control units, respectively. We can see that outcomes are highly correlated within a group.  
```{r fe1, cache=TRUE, message=FALSE, fig.height=4, fig.width=8}
library(ggplot2)
ggplot(s4, aes(x=unit, y = Y, colour = as.factor(D))) + geom_point() + guides(colour=FALSE) 
```

When fixed effects are present, it is possible that we cannot observe a clear pattern of marginal effects in the raw plot as before, while binning estimates have wide confidence intervals:
```{r fe2a, results = "hide", cache=TRUE, message=FALSE, fig.height=4, fig.width=8}
interflex(estimator = "raw", Y = "Y", D = "D", X = "X", data = s4, 
          weights = NULL,ncols=2)
```


```{r fe3, results = "hide", cache=TRUE, message=FALSE, fig.height=5, fig.width=5}
s4.binning <- interflex(Y = "Y", D = "D", X = "X", Z = "Z1", data = s4, 
                        estimator = "binning", FE = NULL, cl = "unit")
plot(s4.binning)
```

The binning estimates are much more informative when fixed effects are included, by using the `FE` option. Note that the number of group indicators can exceed 2. Our algorithm is optimized for a large number of fixed effects or many group indicators. The `cl` option controls the level at which standard errors are clustered. 
```{r fe4, results = "hide", cache=TRUE, message=FALSE, fig.height=5, fig.width=5}
s4$wgt <- 1
s4.binning <- interflex(Y = "Y", D = "D", X = "X", Z = "Z1", data = s4, 
                        estimator = "binning", FE = c("unit", "year"), 
                        cl = "unit", weights = "wgt")
plot(s4.binning)
```

When fixed effects are not taken into account, the kernel estimates are also less precisely estimated. Because the model is incorrectly specified, cross-validated bandwidths also tend to be bigger than optimal. 
```{r fe5, results = "hide", cache=TRUE, message=FALSE, fig.height=5, fig.width=5}
s4.kernel <- interflex(Y = "Y", D = "D", X = "X", Z = "Z1", data = s4, 
                       estimator = "kernel", FE = NULL, 
                       cl = "unit", weights = "wgt")
plot(s4.kernel)
```

Controlling for fixed effects by using the `FE` option solves this problem. The estimates are now much closer to the population truth. Note that all uncertainty estimates are produced via bootstrapping. When the `cl` option is specified, a block bootstrap procedure will be performed.  
```{r fe6, results = "hide", cache=TRUE, message=FALSE, fig.height=5, fig.width=5}
s4.kernel <- interflex(Y = "Y", D = "D", X = "X", Z = "Z1", data = s4, 
                       estimator = "kernel", FE = c("unit","year"), cl = "unit")
plot(s4.kernel)
```

With large datasets, cross-validation or bootstrapping can take a while. One way to check the result quickly is to shut down the bootstrap procedure (using `CI = FALSE`). **interflex** will then present the point estimates only. Another way is to supply a reasonable bandwidth manually by using the `bw` option such that cross-validation will be skipped. 
```{r fe7, results = "hide", cache=TRUE, message=FALSE, fig.height=5, fig.width=5}
s4.kernel <- interflex(Y = "Y", D = "D", X = "X", Z = "Z1", data = s4, 
                       estimator = "kernel", bw = 0.62, FE = c("unit","year"), cl = "unit", 
                       CI = FALSE, ylim = c(-9, 15), theme.bw = TRUE)
plot(s4.kernel)
```

---

## Multiple (>2) treatment arms

Next, we will show how **interflex** can be applied when there are more than two condition. First, we plot the raw data:
```{r mta_raw, results = "hide", cache=TRUE, message=FALSE, fig.height=5, fig.width=10}
interflex(estimator = "raw",Y = "Y", D = "D", X = "X", data = s5, ncols = 3)
```

By default, **interflex** will produce $(n-1)$ marginal effects plots, taking one category as the baseline (if not specified, the base group is selected based on the numeric/character order of the treatment values). Here we also specify `vartype = "bootstrap"` in order to estimate the confidence intervals.
```{r mta1a, results = "hide", cache=TRUE, message=FALSE, fig.height=5, fig.width=8}
s5.binning <- interflex(Y = "Y", D = "D", X = "X", Z = c("Z1","Z2"), data = s5, 
                        estimator = "binning", vartype = "bootstrap")
plot(s5.binning)
```


By setting the option `order` and `subtitles` (in either **interflex** or **plot**), users can also specify the order of plots they prefer and the subtitles they want:  
```{r mta1c, message = FALSE, fig.height=5, fig.width=8, cache = TRUE,results = "hide"}
plot(s5.binning, order=c("C", "B"), subtitles = c("Group C","Group B"))
```

By setting `pool = TRUE` (in either **interflex** or **plot**), we can combine these plots. Users can also specify the colors they like, : 
```{r mta1d, message = FALSE, fig.height=5, fig.width=6, cache = TRUE}
plot(s5.binning, order=c("C", "B"), subtitles = c("Control Group","Group C","Group B"), 
     pool = TRUE, color = c("Salmon","#7FC97F", "#BEAED4"))
```

We specify the base group using the option `base`:
```{r mta2, cache=TRUE, results = "hide", message=FALSE, fig.height=5, fig.width=8}
s5.binning2 <- interflex(Y = "Y", D = "D", X = "X",Z = c("Z1", "Z2"), data = s5, 
                         estimator = "binning", base = "B", vartype = "bootstrap")
plot(s5.binning2)
```


For dataset **s5**, because of the nonlinear interaction effects, the kernel estimator will produce marginal effects estimates that are much closer to the effects implied by the true DGP.
```{r mta3, cache=TRUE, results = "hide", message=FALSE, fig.height=5, fig.width=8}
s5.kernel <- interflex(Y = "Y", D = "D", X = "X",Z = c("Z1", "Z2"), data = s5, 
                       estimator = "kernel")
s5.kernel$figure

plot(s5.kernel, pool = TRUE)
```


---

## Predicted outcomes

Starting from v.1.1.0, **interflex** allows users to estimate and visualize predicted outcomes given fixed values of the treatment and moderator based on flexible interaction models. All predictions are made by setting covariates equal to their sample means. 

```{r ip3, fig.height=5, fig.width=8, cache=TRUE, results = "hide", message=FALSE}
s5.kernel <-interflex(Y = "Y", D = "D", X = "X", Z = c("Z1", "Z2"),data = s5, estimator = "kernel")
predict(s5.kernel)
```

We can also set `pool = TRUE` to combine them.
```{r ip5, fig.height=5, fig.width=8, cache=TRUE, results = "hide", message=FALSE}
predict(s5.kernel,order = c('A','B','C'),subtitle = c("Group A", "Group B", "Group C"), 
        pool = TRUE, legend.title = "Three Different Groups")
```

We can also use the linear model or the binning model to estimate and visualize predicted outcomes. For the latter, it is expected to see some "zigzags" at the bin boundaries in a plot:
```{r ip6, fig.height=5, fig.width=8, cache=TRUE, results = "hide", message=FALSE}
s5.binning <-interflex(Y = "Y", D = "D", X = "X", Z = c("Z1","Z2"), data = s5, 
                       estimator = "binning", vartype = "bootstrap", nbins = 4)
predict(s5.binning)
```

```{r ip7, fig.height=5, fig.width=8, cache=TRUE, results = "hide", message=FALSE}
s5.linear <-interflex(Y = "Y", D = "D", X = "X", Z = c("Z1","Z2"), data = s5, 
                      estimator = "linear", vcov.type = "robust")
predict(s5.linear)
```

Obviously, the kernel estimator characterizes the data more better than a linear or binning estimator.


---

## Differences in treatment effects

Starting from v.1.1.0, **interflex** allows users to compare treatment effects at three specific values of the moderator using linear or kernel model. In order to do so, users need to pass three values into `diff.values` when applying *linear* or *kernel* estimator. If not specified, this programme will by default choose the 25,50,75 percentile of the moderator as the `diff.values`.

```{r dbt1, fig.height=5, fig.width=10, message=FALSE,results = "hide", cache = TRUE}
s5.kernel <-interflex(Y = "Y", D = "D", X = "X", Z = c("Z1", "Z2"), data = s5, 
                      estimator = "kernel", diff.values = c(-2,0,2),
                      vartype = "bootstrap")
plot(s5.kernel,diff.values = c(-2,0,2))
```

```{r dbt2, message=FALSE}
s5.kernel$diff.estimate
```

```{r dbt3, fig.height=5, fig.width=10, message=FALSE,results = "hide", cache = TRUE}
s5.linear <-interflex(Y = "Y", D = "D", X = "X", Z = c("Z1","Z2"), data = s5, 
                      estimator = "linear", diff.values = c(-2,0,2), 
                      vartype = "bootstrap")
plot(s5.linear,diff.values = c(-2,0,2))
```

```{r dbt4, message=FALSE}
s5.linear$diff.estimate
```

Starting from v.1.1.1, **interflex** allows users to compare treatment effects at two or three specific values of the moderator using marginal effects and vcov matrix derived from linear/kernel estimation. Based on GAM model(relies on **mgcv** package), users can approximate treatment effects and their variance using smooth functions without re-estimating the model, hence saving time. In order to do so, users need to pass the output of **interflex** into **t.test** and then specify the values of interest of the moderator. As it is an approximation, the results here are expected to have a little deviance from the output we got by directly specifying `diff.values` when applying **inter.linear** or **inter.kernel**. We also limit the elements passed in diff.values between the minimum and the maximum of the moderator to avoid potential extrapolation bias.

```{r dbt5, message=FALSE, cache = TRUE}
inter.test(s5.kernel,diff.values=c(-2,0,2))
```

```{r dbt6, message=FALSE, cache = TRUE}
inter.test(s5.linear,diff.values=c(-2,0,2))
```

We can also set `percentile = TRUE` to use the percentile scale rather than the real scale of the moderator.
```{r dbt7, message=FALSE, cache = TRUE}
inter.test(s5.kernel,diff.values=c(0.25,0.5,0.75),percentile=TRUE)
```

