# Discrete Moderator {#sec-discrete-moderator}

In this chapter, we explore the special cases of CME where moderator $X$ is discrete. We show how to use **interflex** package for estimation, with binary treatment and continuous treatment.

```{r include=FALSE}
source("init.R")
```

### Binary treatment

When moderator $X$ is discrete, the estimation of CME is a special case of estimating Conditional Average Treatment Effect (CATE), then marginalized over the moderator $X$.

In the **interflex** package, we are estimating CME with discrete moderator $X$ under the Double machine learning (DML) framework. In the settings of binary treatment, what's under the hood is that the data is partitioned into groups based on moderator $X$, within each groups, the function estimate treatment effect separately via cross-fitting, maintaining the orthogonality correction provided by DML.

We use empirical example from @beiser2024pocketbook to illustrate the implementation of **interflex** estimator with binary treatment.

```{r}
d<-load("data/BB2024c6.RData")
d<-as.data.frame(loc_out)
d<-remove_labels(d)

main<-"BB2024c6"
D="treat_type_bin"
Y="sup_tax_rebate" 
X="income"

Z =c("leftright", "clim_belief", "resp_age", "female","employment_f", "highedu", "midedu", "pid_f")
Dlabel<-"Rebate"
Ylabel<-"Carbon Tax Support (with rebate)"
Xlabel<-"Income"
```

**Visualization**

```{r}
out.raw <- interflex(estimator='raw', data = d, 
                       Y = Y, D = D, X = X, Z = Z, na.rm = TRUE,
                       Ylabel = Ylabel, Dlabel = Dlabel,
                       Xlabel = Xlabel, theme.bw = TRUE)

out.raw
```

From the scatter plot, the treatment effect is positively correlated to moderator value $[5,8]$.

**Estimation.** To estimate CME, we specify the `estimator = 'dml'` and set the option `gate = TRUE`.

```{r}
out.gate <- interflex(estimator='dml', data = d, 
                       Y = Y, D = D, X = X, Z = Z, na.rm = TRUE,
                       Ylabel = Ylabel, Dlabel = Dlabel,
                       Xlabel = Xlabel, theme.bw = TRUE,
                       gate = TRUE)
out.gate$figure
```

The default visualization gives a smooth line even though our moderator is discrete. Each position on the x‐axis is just one of the discrete income categories. The black curve is the estimated average treatment effect for each "group" (each category), and the gray ribbon and dash line are the confidence interval.

For a more clear and accurate visualization, we can use the `plot()` function and set `by.group = TRUE`. The figure shows a separate point estimate and confidence interval, both pointwise CI, in solid black, and uniform CI, in shallow gray, for each discrete level of the moderator. The plot show an obvious positive CME at $X = 7$, but the treatment effect is not statistically significant.

```{r}
plot(out.gate, by.group = TRUE, theme.bw = TRUE)
```

**Results.** User can extract the estimation by group through `g.est.dml`.

```{r}
out.gate$g.est.dml[[1]]
```

------------------------------------------------------------------------

### Continuous treatment treatment

For CME with discrete moderator and continuous treatment, we are using a simulated data with treatment $D$ following the normal distribution, a discrete moderator with values ${1,2,3,4,5}$, assigned with heterogeneous effect, and two covariates $Z_1, Z_2$. 

```{r}
set.seed(9999)
n <- 2000

D <- rnorm(n, mean = 3, sd = 1)
X <- sample(1:5, size = n, replace = TRUE, prob = c(0.1, 0.3, 0.3, 0.2, 0.1))

## covariates
Z1 <- runif(n, min = -2, max = 2)
Z2 <- rgamma(n, shape = 2, rate = 1)

slopes_by_x <- c(`1`=2, `2`=4, `3`=7, `4`=3, `5`=10)
tau_i <- slopes_by_x[as.character(X)] + rnorm(n, mean = 0, sd = 0.5)

alpha <- 5

beta1 <- 1.5    # coefficient on Z1^2
beta2 <- -2.0   # coefficient on sqrt(Z2)

e <- rnorm(n, mean = 0, sd = 2)

## Y = alpha + tau_i * D + beta1*(Z1^2) + beta2*sqrt(Z2) + e
Y <- alpha + tau_i*D + beta1*(Z1^2) + beta2*sqrt(Z2) + e
s10 <- data.frame(Y, D, X, Z1, Z2)

head(s10)
```


The syntax to estimate CME with continuous treatment and discrete moderator is as follow. User can plot the treatment effect based on individual discrete categories with separate point estimate and confidence interval. 
```{r}

out.gate.con <- interflex(estimator='dml', data = s10, 
                       Y = "Y", D = "D", X = "X", Z = c("Z1", "Z2"),
                       theme.bw = TRUE,
                       gate = TRUE)
out.gate.con$figure

plot(out.gate.con, by.group = TRUE)

```



