---
title: 'Tutorial: Interflex with Double Machine Learning'
output:
  rmarkdown::html_vignette:
    toc: false
    toc_depth: 4
    number_sections: false
vignette: >
  %\VignetteIndexEntry{Tutorial: Interflex with Double Machine Learning}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}  
---
<!-- 
  Code to Justify Text
    <style>
    body {
    text-align: justify}
    </style>
-->   
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Load the package as well as the simulated toy datasets for this tutorial. **s5** is for the tutorial for **interflex** with continuous outcomes.

```{r, message = FALSE}
library(interflex)
data(interflex)

set.seed(1234)
n<-200
x<-rnorm(n,3,1) # moderator
z<-rnorm(n,3,1) # covariate
e<-rnorm(n,0,1) # error term

## Multiple treatment arms
n <- 600
# treatment 1
d1 <- sample(c('A','B','C'),n,replace=T)
# moderator
x <- runif(n,min=-3, max = 3)
# covriates
z1 <- rnorm(n,0,3)
z2 <- rnorm(n,0,2)
# error
e <- rnorm(n,0,1)
y1 <- rep(NA,n)
y1[which(d1=='A')] <- -x[which(d1=='A')]
y1[which(d1=='B')] <- (1+x)[which(d1=='B')]
y1[which(d1=='C')] <- (4-x*x-x)[which(d1=='C')]
y1 <- y1 + e + z1 + z2
s5 <- cbind.data.frame(D=d1, X=x, Y=y1, Z1 = z1,Z2 = z2)
```

---

## Double Machine Learning

When setting the option `estimator` to "DML" and the option `ml_method` to "nn", the program will conduct the two-stage estimator (with machine learning method as Neural Network) to compute the marginal effects/treatment effects. In the following command, the outcome variable is "Y", the moderator is "X", the covariates are "Z1" and "Z2", and here it uses logit link function.
```{r dlinear1, cache=TRUE, results = "hide", message=FALSE}
s5.DML.nn <- interflex(estimator='DML', data = s5, ml_method="nn",
                       Y = "Y", D = "D", X = "X", Z = "Z1", treat.type = "discrete")
```

### Marginal/treatment effects

The estimated treatment effects of "D" on "Y" across the support of "X" are saved in 
```{r dlinear2, cache=TRUE}
s5.DML.nn$est.dml
```

Users can then use the command **plot** to visualize the treatment effects
```{r dlinear3, results = "hide", fig.height=5, fig.width=8}
s5.DML.nn.plot <- plot(s5.DML.nn)
s5.DML.nn.plot
```

### Parameters

Users can change the parameters for the machine learning method in Stage 1. 

```{r dlinear4, results = "hide", fig.height=5, fig.width=8}
s5.DML.nn.2 <- interflex(estimator='DML', data = s5, ml_method="nn",
                       Y = "Y", D = "D", X = "X", Z = c("Z1", "Z2"),
                       solver = "lbfgs",
                       max_iter = 10000,
                       alpha = 1e-5,
                       hidden_layer_sizes = c(3, 3, 2),
                       random_state = 1,
                       treat.type = "discrete")
plot(s5.DML.nn.2)
```

### Machine Learning Methods

Users can also change the machine learning method Stage 1. Now `Random Forest` (rf), `Hist Gradient Boosting` (hgb), and `Neural Network` (nn) are allowed to input as `ml_method`.

```{r dlinear5, results = "hide", fig.height=5, fig.width=8}
s5.DML.rf <- interflex(estimator='DML', data = s5, ml_method="RF",
                       Y = "Y", D = "D", X = "X", Z = c("Z1", "Z2"), treat.type = "discrete")
plot(s5.DML.rf)

s5.DML.hgb <- interflex(estimator='DML', data = s5, ml_method="hgb",
                    Y = "Y", D = "D", X = "X", Z = c("Z1", "Z2"), treat.type = "discrete")
plot(s5.DML.hgb)
```


