# Get Started {#sec-introduction}

## Installation

To use the **interflex** package, please install directly from Github:

```{r, eval=FALSE}

# if not already installed
install.packages('devtools', repos = 'http://cran.us.r-project.org') 
devtools::install_github('xuyiqing/interflex@dml', force = TRUE)
```

### Python environment setup

To use estimators in **interflex**, we rely on several Python packages. The integration between R and Python is facilitated by the **reticulate** package, which allows R to interface with Python seamlessly. The following steps set up a Python environment with all required dependencies.

First, install the **reticulate** package in R. This package enables R to interface with Python.

```{r, eval=FALSE}
install.packages("reticulate", repos = 'http://cran.us.r-project.org')
```

Then, we use [Miniconda](https://docs.anaconda.com/free/miniconda/), a minimal installer for Conda. Miniconda helps manage Python environments and packages. By default, the installation process will create a virtual environment named `r-reticulate`.

```{r, eval=FALSE}
reticulate::install_miniconda(update = TRUE)
```

With Miniconda installed, we point to the platform using `use_condaenv`. This step is incorporated in the **interflex** package.

```{r, eval=FALSE}
reticulate::use_condaenv(condaenv = "r-reticulate")
```

Finally, we install the necessary Python libraries. These libraries include tools for statistical modeling and machine learning that are essential for using the **interflex** package.

```{r, eval=FALSE}
reticulate::py_install(packages = c("patsy", "numpy", "pandas",
                                    "scikit-learn", "doubleml", "econml"))
```

Some Mac users may encounter a “\*\*\* caught segfault \*\*\*” error. In that case, we recommend setting up a virtual environment. In Terminal, type:

```         
python3 -m venv ~/.virtualenvs/myenv source ~/.virtualenvs/myenv/bin/activate
```

Then install all Python packages there:

```         
pip install numpy
pip install patsy
pip install pandas
pip install scikit-learn
pip install econml
pip install doubleml
```

In R, point to the newly set up environment `myenv`:

```{r, eval=FALSE}
reticulate::use_virtualenv("myenv")
```

### **Load packages**

**interflex** depends on the following packages, which will be installed automatically when **interflex** is being installed; if not, please install them manually:

```{r eval=FALSE, message=FALSE}
# Function to install packages
install_all <- function(packages) {
  installed_pkgs <- installed.packages()[, "Package"]
  for (pkg in packages) {
    if (!pkg %in% installed_pkgs) {
      install.packages(pkg, repos = 'http://cran.us.r-project.org')
    }
    library(pkg, character.only = TRUE)
  }
}

# Packages to be installed
packages <- c("Rcpp", "mgcv", "sandwich", "pcse", "fixest", "foreach",
              "doParallel", "lfe", "lmtest", "Lmoments", "ggplot2", 
              "plotrix", "grid", "gridExtra", "ggplotify", "ggpubr",
              "RColorBrewer", "grDevices", "gtable", "MASS", "mvtnorm", 
              "pROC", "ModelMetrics", "foreign","patchwork", "rmarkdown", "DT")

# Install the packages
install_all(packages)

```

## Datasets

The **interflex** package incorporates several datasets:

```{r}
library(interflex)
data(interflex)
ls()
```

-   `s1-s9`: simulation data
-   `app_hma2015`: Data from Huddy 2015, where they study the effect of expressive model of partisanship.


## Set up, estimand, and identification assumption

Understanding conditional relationships is central to social science. Researchers are particularly interested in how the effect of a treatment \( D \) on an outcome \( Y \) changes with the value of a pre-determined covariate \( X \), known as the moderator, which is unaffected by \( D \). research. To formalize this problem better, we introduce the setup and estimand using the Neyman-Rubin potential outcome framework and notation. 

Consider a study with $n$ units, indexed by $i = 1, 2, \dots, n$. These units are independently and identically distributed (i.i.d.) from an infinite population denoted by $\mathcal{P}$, the super-population. 
For each unit $i$, we observe the quadruple $(Y_i, D_{i}, X_i, Z_i)$, where $Y_i$ denotes the outcome of interest, $D_{i}$ is the treatment indicator, and $V_i = (X_i, Z_i)$ represents a set of covariates not causally affected by the treatment and comprises two components: the moderator of interest $X_i \in \mathcal{X}$, which can be either discrete or continuous, and the remaining covariates used for identification $Z_i \in \mathcal{Z}$, which can be potentially high-dimensional. 

Given a sets of assumption, including random sampling, stable unit treatment value assumption (SUTVA), we define Conditional Marginal Effect (CME) with binary treatment as: 

\begin{equation*}
\theta(x) = \mathbb{E}[Y_i(1) - Y_i(0) \mid X_i = x].
\end{equation*} 

And CME with continuous treatment: 

\begin{equation*}  
\theta(x)\;=\;\mathbb{E}\!\big[\,\mu_d(D,X,Z)\ \big|\ X=x\big].
\end{equation*} 
where the dose–response function $\mu(d,x,z) = \mathbb{E}\left[Y(d) \mid X = x, Z = z\right]$ describes how the expected potential outcome varies with the treatment level $d$, conditional on the covariates. The CME, $\theta(x)$, averages the unit-specific rates of change across units with the same value of the moderator $x$. 


We further assume Unconfoundedness and Strict overlap as identification assumptions, which turn CME from causal estimand to statistical estimand for estimation in the coming chapters. 


