# Get Started {#sec-introduction}

## Installation

To install **interflex** from CRAN, run the code chunk below:

```{r install-cran, eval=FALSE, message=FALSE, warning=FALSE}
install.packages("interflex")
```

We recommend users to install the most up-to-date, stable version of **interflex** from GitHub using:

```{r install-github-main, eval=FALSE, message=FALSE, warning=FALSE}
devtools::install_github("xuyiqing/interflex")
```

We fix bugs in the `dev` branch before merging into the main branch; therefore, it is often more up to date.

```{r install-github-dev, eval=FALSE, message=FALSE, warning=FALSE}
devtools::install_github("xuyiqing/interflex@dev")
```

After installation, check **interflex** version to make sure the package is up-to-date.

```{r check-version}
installed.packages()["interflex", "Version"]
```

### Optional GitHub-only dependencies

**interflex** optionally integrates with the **mlr3extralearners** package for additional machine learning backends. **mlr3extralearners** is not on CRAN but can be installed from GitHub:

```{r install-mlr3extra, eval=FALSE, message=FALSE, warning=FALSE}
devtools::install_github("mlr-org/mlr3extralearners")
```

### Dependencies

**interflex** depends on the following packages, which should be installed automatically when **interflex** is being installed. You can also install them manually.

```{r install-dependencies, eval=FALSE}
install_all <- function(packages) {
  installed_pkgs <- installed.packages()[, "Package"]
  for (pkg in packages) {
    if (!pkg %in% installed_pkgs) {
      install.packages(pkg)
    }
  }
}
packages <- c("Rcpp", "RcppArmadillo", "ggplot2", "sandwich", "Lmoments",
              "doParallel", "foreach", "mgcv", "lfe", "gridExtra", "grid",
              "ggplotify", "fixest", "RColorBrewer", "pcse", "gtable",
              "MASS", "mvtnorm", "pROC", "ModelMetrics", "lmtest", "AER",
              "future", "glmnet", "grf",
              "DoubleML", "mlr3", "mlr3learners", "data.table", "paradox", "ranger")
install_all(packages)
```

## Datasets

The **interflex** package incorporates several datasets:

```{r}
library(interflex)
data(interflex)
ls()
```

-   `s1-s9`: simulation data
-   `app_hma2015`: Data from Huddy 2015, where they study the effect of expressive model of partisanship.


## Set up, estimand, and identification assumption

Understanding conditional relationships is central to social science. Researchers are particularly interested in how the effect of a treatment \( D \) on an outcome \( Y \) changes with the value of a pre-determined covariate \( X \), known as the moderator, which is unaffected by \( D \). research. To formalize this problem better, we introduce the setup and estimand using the Neyman-Rubin potential outcome framework and notation. 

Consider a study with $n$ units, indexed by $i = 1, 2, \dots, n$. These units are independently and identically distributed (i.i.d.) from an infinite population denoted by $\mathcal{P}$, the super-population. 
For each unit $i$, we observe the quadruple $(Y_i, D_{i}, X_i, Z_i)$, where $Y_i$ denotes the outcome of interest, $D_{i}$ is the treatment indicator, and $V_i = (X_i, Z_i)$ represents a set of covariates not causally affected by the treatment and comprises two components: the moderator of interest $X_i \in \mathcal{X}$, which can be either discrete or continuous, and the remaining covariates used for identification $Z_i \in \mathcal{Z}$, which can be potentially high-dimensional. 

Given a sets of assumption, including random sampling, stable unit treatment value assumption (SUTVA), we define Conditional Marginal Effect (CME) with binary treatment as: 

\begin{equation*}
\theta(x) = \mathbb{E}[Y_i(1) - Y_i(0) \mid X_i = x].
\end{equation*} 

And CME with continuous treatment: 

\begin{equation*}  
\theta(x)\;=\;\mathbb{E}\!\big[\,\mu_d(D,X,Z)\ \big|\ X=x\big].
\end{equation*} 
where the dose–response function $\mu(d,x,z) = \mathbb{E}\left[Y(d) \mid X = x, Z = z\right]$ describes how the expected potential outcome varies with the treatment level $d$, conditional on the covariates. The CME, $\theta(x)$, averages the unit-specific rates of change across units with the same value of the moderator $x$. 


We further assume Unconfoundedness and Strict overlap as identification assumptions, which turn CME from causal estimand to statistical estimand for estimation in the coming chapters. 


